---
title: "ICCONIC HipFX time to surgery: Analytical pipeline"
author: Data Science for Health Services and Policy Research (IACS)
editor: source

#date: 
# bibliography: 
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 7
    highlight-style: pygments
    code-fold: true
    html-math-method: katex
execute: 
  warning: false
  cache: false
  echo: false
params:
    data_path: '../../inputs/data.duckdb'
    state: ''
---

```{css, echo = FALSE}
.justify {
  text-align: justify !important
}
```


```{r, output = FALSE, echo=FALSE, warning=FALSE}
#| label: load r packages
library(duckdb)
library(dplyr)
library(Hmisc)
library(survival)
library(survminer)
library(lubridate)
library(mgcv)
library(logger)
library(stringr)
library(sjPlot)
library(gt)
library(reshape2)
library(htmltools)
library(ggridges)
library(performance)
# library(ggplotify)

#new
library(lme4)
library(gratia)

s <- Sys.getenv("PIPELINE_VERSION")

log_info(paste0("Analysis version: ",s))

```

<p> Analysis version: `r paste0(s)` </p>



```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: functions
calculate_MOR <- function(model,var_re,n_iter){
  #calculate MOR with 95% CI
  if(inherits(model,'glmerMod')){
    t <- lme4::ranef(model,condVar = TRUE)
    est<-as.numeric(unlist(t[[var_re]]))
    
    var<- as.numeric(unlist(attr(t[[var_re]],"postVar")))
  }else{
    idx <-grep(var_re, names(coef(model)))
    est<-coef(model)[idx]
    attributes(est)<-NULL
    var <-sqrt(diag(model$Ve))[idx]
    }
    
  ### bootstrap 
  ### create empty output collection:
  mor_boot<- c()
  #### iterate over replicates
  for(i in 1:n_iter){
    ### draw vector of area random effects from normal
    drw<- rnorm(n = length(est),mean = est,sd = sqrt(var))
    ### create data frame with all possible pairs 
    s<- combn(drw,2)
    #### estimate MOR and save in output
    mor_boot<- c(mor_boot, median(exp(abs(s[1,]- s[2,]))))
  }
  ### bootstrap median and 95% CI
  p <- quantile(mor_boot, c(.025,.5,.975))
  return(p)
}

calculate_ICC <- function(model, variable_random_effect){
  if(inherits(model,'glmerMod')){
    x <- summary(model)$varcor
    x <- as.data.frame(x)
    icc <- (x$vcov)/((x$vcov)+((pi*pi)/3))  
    icc_lower <- (x$vcov - x$sdcor)/((x$vcov - x$sdcor)+((pi*pi)/3))
    icc_upper <- (x$vcov + x$sdcor)/((x$vcov + x$sdcor)+((pi*pi)/3))
    print(paste0('ICC ajusted: ', round(icc,5),' with lower_ci95=',round(icc_lower,5),' and upper_ci95=',round(icc_upper,5)))
    return(x)
  }else{
    var <- gratia::variance_comp(model)
    
    print(var)
    var_re <- var %>% filter(grepl(variable_random_effect, component))
    icc <- (var_re$variance/(var_re$variance + ((pi*pi)/3)))
    icc_lower <-  (var_re$variance - var_re$std_dev)/((var_re$variance - var_re$std_dev)+((pi*pi)/3))
    icc_upper <-  (var_re$variance + var_re$std_dev)/((var_re$variance + var_re$std_dev)+((pi*pi)/3))
    print(paste0('ICC ajusted: ', round(icc,5),' with lower_ci95=',round(icc_lower,5),' and upper_ci95=',round(icc_upper,5)))
    return(var_re)
  }
}

```


```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: query descriptive patient


con = dbConnect(duckdb::duckdb(), dbdir=params$data_path, read_only=FALSE)

descriptive_values <- dbGetQuery(conn = con, "
SELECT first (country_cd) filter (where country_cd IS NOT NULL) as country_cd,
median(age_nm) as median_age,
(QUANTILE_CONT(age_nm, 0.75)-QUANTILE_CONT(age_nm, 0.25)) as iqr_age,
count(distinct patient_id) filter (where sex_cd = '1') as men,
count(distinct patient_id) filter (where sex_cd = '2') as women,
count(distinct patient_id) as n_patient,
count(distinct hospitalization_episode_id) as n_hospitalization_episode_id,
count(distinct patient_id) filter (where surgical_procedure_bl = TRUE) as n_patient_with_surgery,
median(surgical_procedure_dt -hospital_admission_dt) as median_time_adm_to_surgery,
(QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.75)-QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.25)) as IQR_median_time_adm_to_surgery,
-- round(mean(surgical_procedure_dt -hospital_admission_dt),2) as mean_time_adm_to_surgery
FROM main.patient_cohort")


descriptive_values <- descriptive_values %>% mutate(
  perc_men = round((men/n_patient)*100,2),
  united_men =paste0(men, ' (',perc_men,'%)' ),
  perc_women = round((women/n_patient)*100,2),
  united_women =paste0(women, ' (',perc_women,'%)' )
)

descriptive_values <- descriptive_values %>% select(!c(men,women,perc_men,perc_women))
## socec

socecon_lvl <- dbGetQuery(conn = con,"
SELECT
	coalesce(socecon_lvl_cd, 'Unknown') as socecon_lvl_cd,
	count(distinct patient_id) as patient_distinct_socecon,
	
FROM
	main.patient_cohort
GROUP BY
	socecon_lvl_cd ")
socecon_lvl <- socecon_lvl %>% mutate(
  perc_socecon_lvl = round((patient_distinct_socecon/descriptive_values$n_patient)*100,2),
  united_socecon =paste0(patient_distinct_socecon, ' (',perc_socecon_lvl,'%)' )
)

## surgical_procedure_cd
surgical_procedure <- dbGetQuery(conn = con,"
SELECT
	coalesce(surgical_procedure_cd::int,
	'Non surgical procedure' ) as surgical_procedure_cd ,
	count(distinct patient_id) as patient_distinct_surgical_procedure_cd
FROM
	main.patient_cohort
GROUP BY
	surgical_procedure_cd")
	
surgical_procedure <- surgical_procedure %>% mutate(
  perc_surgical_procedure = round((patient_distinct_surgical_procedure_cd/sum(surgical_procedure$patient_distinct_surgical_procedure_cd))*100,2),
  united_surgical_procedure=paste0(patient_distinct_surgical_procedure_cd, ' (',perc_surgical_procedure,'%)' )
)

## anesthesia_type_cd only in patients with surgical procedure, so perc is divided by number of patients with surgery

anesthesia_type <- dbGetQuery(conn=con,"
 SELECT
	coalesce(anesthesia_type_cd::int,
	'Unknown') as anesthesia_type_cd,
	count(distinct patient_id) as patient_distinct_anesthesia_type_cd
FROM
	main.patient_cohort
WHERE surgical_procedure_bl = TRUE
GROUP BY
	anesthesia_type_cd ")

anesthesia_type <- anesthesia_type %>% mutate(
  perc_anesthesia_type = round((patient_distinct_anesthesia_type_cd/descriptive_values$n_patient_with_surgery)*100,2),
  united_anesthesia_type=paste0(patient_distinct_anesthesia_type_cd, ' (',perc_anesthesia_type,'%)' )
)


comorbidities <- dbGetQuery(conn=con,"
with seleccion as (
SELECT
	patient_id ,
	list_aggregate([chronic_kidney_disease_bl::int,
	tobacco_copd_bl::int,
	depression_bl::int,
	serious_mental_illness_bl::int,
	alcohol_abuse_bl::int,
	obesity_overweight_bl::int,
	osteoporosis_osteopenia_bl::int,
	dementia_bl::int,
	diabetes_bl::int,
	liver_disease_bl::int,
	pancreatic_disease_bl::int,
	inflamatory_bowel_disease_bl::int,
	rheumatology_disease_bl::int,
	spinal_cord_disease_injury_bl::int,
	serious_neurological_disease_bl::int,
	parkinson_huntington_diseases_bl::int,
	seizure_disorder_convulsions_bl::int,
	congestive_heart_failure_bl::int,
	coronary_artery_disease_bl::int,
	cerebrovascular_disease_bl::int,
	peripheral_vascular_disease_bl::int,
	traumatic_brain_injury_bl::int,
	amputee_bl::int],
	'sum') as sum_comorbidities
from
	main.patient_cohort)
select
	sum_comorbidities,
	count(DISTINCT patient_id) as patient_distinct_n_comorbidities
from
	seleccion
group by
	sum_comorbidities")

comorbidities_ <- data.frame(comor_0_2 = sum(filter(comorbidities, sum_comorbidities <= 2)$patient_distinct_n_comorbidities),
                             comor_3_5 = sum(filter(comorbidities, sum_comorbidities > 2 & sum_comorbidities <=5)$patient_distinct_n_comorbidities),
                             comor_5mayor = sum(filter(comorbidities,sum_comorbidities >5)$patient_distinct_n_comorbidities))
comorbidities_ <- melt(comorbidities_)
comorbidities_ <- comorbidities_ %>% mutate(
  perc_comorbidities = round((value/descriptive_values$n_patient)*100,2),
  united_comorbidities=paste0(value, ' (',perc_comorbidities,'%)' )
)
rm(comorbidities)
```


```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: query descriptive hospital

variables_in_hospital <- dbGetQuery(conn=con,"
SELECT
	coalesce(hospital_st,
	'Unknown')  as hospital_st,
count(distinct patient_id) as n_patient,
count(distinct hospitalization_episode_id) as n_hospitalization_episode_id,
count(distinct patient_id) filter (where surgical_procedure_bl = TRUE) as n_patient_with_surgery,
median(age_nm) as median_age, 
(QUANTILE_CONT(age_nm, 0.75)-QUANTILE_CONT(age_nm, 0.25)) as iqr_age,
count(distinct patient_id) filter (where sex_cd = '2') as women,
median(surgical_procedure_dt -hospital_admission_dt) as median_time_adm_to_surgery,
(QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.75)-QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.25)) as IQR_median_time_adm_to_surgery,
count(distinct patient_id) filter (where exitus_bl = TRUE) as n_exitus,
count(distinct patient_id) filter (where unplanned_readmission_30d_bl = TRUE) as n_unplanned_readmission_30d,
from
	main.patient_cohort
group by
	hospital_st")

dbDisconnect(con, shutdown=TRUE)

```

# Introduction

::: {.justify}
Breaking a hip might be a sign of frailty. It can entail diminished autonomy and a high risk of death in the elderly (i.e. 30% dying within a year, with 7% dying during hospitalization). With the ageing population, hip fractures in the elderly are becoming a growing medical and social concern as it lead to frailty, decreasing autonomy, worse quality of life or even death. 

In hip fracture care, whether mortality worsens when surgery is delayed is disputed. Proponents of early surgery believe that there is decreased morbidity with early fixation and reduced mortality with early mobilization postoperatively. Proponents of elective surgery argue that most elderly require vascular stabilization, including reversion of antiplatelet or anticoagulant therapies to minimize bleeding during hip surgery, resulting in better quality of care and, ultimately, better outcomes.    

Key elements of the acute treatment of hip fracture in the elderly (frailty) population imply health care capacity and processes in weighting the individual situation of each patient in terms of prognosis, concurrent clinical conditions, medication, autonomy and self-care ability, and social circumstances in the indication of surgery with the expectation for the patient returning home. 
Mortality in hip fracture patients after surgery can be a critical indicator in benchmarking hospital performance in assisting the elderly and frail population.  

An exploratory analysis of the prevalence of hip fracture surgery (i.e., hip pinning, hip replacement and hip repair) in the frail population was conducted in several countries participating in ICCONIC as part of the project's first phase. Time to hip fracture surgery, a classical predictor of good hip fracture repair outcomes, varied across ICCONIC countries. Likewise, preliminary results exhibited large differences in health outcomes, particularly in-hospital and 30-day mortality. However, no substantial differences in mortality were observed among the ICCONIC countries, with the largest difference in time to surgery (Germany and Spain).

This preliminary evidence is consistent with literature focusing on the type of surgery and the management of anti-aggregated or anticoagulated patients - rather than on time-to-surgery - as the main predictors of adverse outcomes.

However, the evidence from ICCONIC is based on aggregated information that might suffer from some biases. Taking advantage of the ICCONIC collaborative at (https://www.lse.ac.uk/lse-health/research/projects/ICCONIC) and (https://icconic.org/), we will explore the variation in inpatient, 7-day mortality and 30-day mortality after hip fracture surgery across ICCONIC hospitals, using individual-level data, taking into consideration the differences in surgical and perioperative management as well as health system differences. In addition, we will examine patterns of polypharmacy, high-risk prescribing, and medication review before and after hip fracture. Subsequently, we will determine hip fracture outcomes depending on socioeconomic factors.
This study aims to measure mortality after hip fracture surgery as a benchmark for hospital performance across several countries, considering patients’, hospitals’ and contextual factors that might contribute to better outcomes, including time-to-surgery, comorbidities and previous therapies.

:::
# Hypothesis
::: {.justify}
The study's central hypothesis explores whether hospital performance in terms of mortality after surgery worsens when the intervention is delayed after a hip fracture in a population over 65. 
:::
# Objectives

## General objective
::: {.justify}
Determinate that prompt action in the acute treatment of hip fracture in the elderly (frailty) population implies health care capacity and processes in weighting the individual situation of each patient in terms of prognosis, concurrent clinical conditions, medication, autonomy and self-care ability, and social circumstances in the indication of surgery with the expectation for the patient returning home.
:::
### Specific Objectives
::: {.justify}
1.	Evaluate the impact of time to hip surgery after fracture on mortality 

2.	Determinate hip fracture outcomes depending on socioeconomic factors
:::
# Methodology

## Design
::: {.justify}
An observational retrospective longitudinal multi-country cohort study of individuals aged 65+ admitted to a hospital due to hip fracture that underwent hip pinning or replacement surgery between 01-01-2017 and 31-12-2021.  
The cohort to be studied includes any person 65 years old or older, with an unplanned hospital episode (inpatient), with a primary diagnosis of hip fracture, with the following exclusion criteria: a) any person under 65 years old, b) multiple fractures (polytrauma), or c) admission from another hospital.
To determine hip fracture outcomes depending on socioeconomic factors would be supplemented with analyses by socio-economic status.
The study will be conducted simultaneously in Spain (Aragón), the USA (Brown University), and several other countries (i.e., Denmark, Finland, Canada, and Sweden), following a federated research methodology based mainly on the reuse of healthcare data from administrative databases at regional or population level at each country.
:::

## Information sources

::: {.justify}
The analyses are based on Individual-level data from health records (i.e., electronic medical records) and administrative data (information on hospital discharges). Data will be analysed locally by each of the participants using deployable analytical scripts. The participants will then share aggregated outputs of the local analyses to conduct a meta-analysis comparing hospitals’ performance in the different countries. 
In Aragon, data will be requested from BIGAN (established by Executive Order SAN/1355/2018 as an element of the health information system in Aragon).
Each participant will comply with their institution's data application processes to access and process data for research. 
::: 

## Population studied
::: {.justify}
All the population insured in Aragon with a relevant contact with the healthcare system during the study period.
:::
## Relevant outcomes
::: {.justify}
*Adjusted inpatient mortality by hospital:* adjusted mortality after surgery due to hip fracture occurring in the index hospital episode considering patient attributes (i.e., age, sex, comorbidities, etc.), care process (i.e. time to surgery, month, weekday, etc.), and contextual data (i.e., socioeconomic data of the area of residence of the patient, etc.)

*7-day adjusted mortality:* adjusted mortality up to 7 days after surgery due to hip fracture considering patient attributes (i.e., age, sex, comorbidities, etc.), care process (i.e. time to surgery, month, weekday, etc.), and contextual data (i.e., socioeconomic data of the area of residence of the patient, etc.)

*30-day adjusted mortality:* adjusted mortality up to 30 days after surgery due to hip fracture considering patient attributes (i.e., age, sex, comorbidities, etc.), care process (i.e. time to surgery, month, weekday, etc.), and contextual data (i.e., socioeconomic data of the area of residence of the patient, etc.)

*7-day unplanned readmission:* adjusted unplanned readmission (for any cause) up to 7 days after hospitalization due to hip fracture considering patient attributes (i.e., age, sex, comorbidities, etc.), care process (i.e. time to surgery, month, weekday, etc.), and contextual data (i.e., socioeconomic data of the area of residence of the patient, etc.)

*30-day unplanned readmission:* adjusted unplanned readmission (for any cause) up to 30 days after hospitalization due to hip fracture considering patient attributes (i.e., age, sex, comorbidities, etc.), care process (i.e. time to surgery, month, weekday, etc.), and contextual data (i.e., socioeconomic data of the area of residence of the patient, etc.)

:::
## Variables of the study
::: {.justify}
You can see in **docs/CDM** folder the specification of the common data model and cohort selection criteria, including the description of the study's variables.
:::
## Statistical Analysis

### Data quality analysis on local inputs
::: {.justify}
Local analyses will be preceded by a data quality and validation analysis on local inputs at the origin. Those analyses will consist of univariate exploratory data analysis of each of the variables of the data entities required for the study and information on the compliance of the local input data with the data validation rules established by the common data model specification. 
:::
### Descriptive statistics
::: {.justify}
Local analysis will include descriptive statistics at the individual level on patient and episode characteristics and prescriptions before and after hospital admission due to hip fracture and on the characteristics of patients admitted and intervened in each hospital. 
:::
### Modelling hospital performance

#### Adjusted mortality by time to surgery
::: {.justify}
In order to study whether there is a relationship between mortality and the time elapsed between the date of admission and the date of surgery, two different groups of models depending on the number of independent variables are carried out. On the one hand, several reference univariate models. On the other hand, several multivariate models.
:::

##### Adjusted mortality by time to surgery (univariate models)
::: {.justify}
Four different univariate models are built:

1. Generalized Linear Model (GLM): three GLM models following the same methodology but varying the dependent variable (y):

    - Model 1: y = inpatient mortality due to any cause

    - Model 2: y = mortality 7-days after surgery (any cause)

    - Model 3: y = mortality 30-days after surgery (any cause)
  
  In this model, the independent variable (x) is the time elapsed between the date of admission and the date of surgery categorized as:
  
    - 1: x <= 24 hours 

    - 2: 24 hours < x <= 48 hours

    - 3: 48 hours < x <= 72 hours

    - 4: 72 hours < x
  

2. Generalized Linear Mixed Model (GLMM): a single GLMM model is constructed following the GLM model methodology, but only for the independent variable *y = mortality 30-days after surgery (any cause)* and including the hospital as the second level.

3. Generalized Additive Model (GAM): three GAM models following the same methodology but varying the dependent variable (y) as in the GLM models, but in this case, the dependent variable is not categorized but introduced as a spline.

4. Generalized Additive Mixed Model (GAMM): a single GAMM model is constructed following the GAM model methodology, but only for the independent variable *y = mortality 30-days after surgery (any cause)* and including the hospital as the second level.
:::

##### Adjusted mortality by time to surgery (multivariate models)
::: {.justify}
Three multilevel GAMM models are built depending on the dependent variable to be predicted (y) as previous section. The most relevant independent variable in this case is (as previous section) the time elapsed between the date of admission and the date of surgery but in this case different patient variables that are considered relevant in the study such as age, sex, comorbidities etc. are introduced into the model.  
:::

#### Adjusted unplanned readmissions by time to surgery (multivariate models)
::: {.justify}
In order to study whether there is a relationship between unplanned readmissions and the time elapsed between the date of admission and the date of surgery, two multivariate GAMM models are carried out

    - Model 1: y = unplanned readmission (for any cause) in the next 7 days after hospitalization

    - Model 2: y = unplanned readmission (for any cause) in the next 30 days after hospitalization

:::
### International hospital performance comparison (meta-analysis)
::: {.justify}
A meta-regression will be adjusted considering region or country-level contextual factors using the aggregated local output data on hospital performance as the primary input. The international comparison meta-analysis will compare hospital performance across the country/regions participating in the study, enabling ranking and providing information on the overall association of healthcare behaviour, namely time-to-surgery on mortality after surgery for hip fracture. 
:::


# Results  (`r descriptive_values$country_cd`)

## Descriptive statistics
::: {.justify}
The following are shown 2 descriptive tables of the imported data:

- Table 1 shows the main characteristics of the patients registered in the database.

- Table 2 main characteristics of the patients and the medians from admission to surgery (in days) and their interquartile range grouped by hospital.

:::
### Table 1. Patient-level summary 

```{r, echo=FALSE,warning=FALSE}
#| label: descriptive tables inputs1
descriptive_values_ <- melt(descriptive_values,id=NULL) 

anesthesia_type <- anesthesia_type %>% arrange(anesthesia_type_cd) %>%  dplyr::select(anesthesia_type_cd,united_anesthesia_type)
anesthesia_type <- anesthesia_type %>% mutate(
    descrp = case_when(
    anesthesia_type_cd %in% '0' ~ 'General',
    anesthesia_type_cd %in% '1' ~ 'Locorregional',
    anesthesia_type_cd %in% '2' ~ 'Locorregional and General',
    TRUE ~ 'Unknown'
    )
)


comorbidities_ <- comorbidities_ %>% arrange(variable) %>% dplyr::select(variable,united_comorbidities) %>% 
  mutate(descrp = case_when(
    variable %in% 'comor_0_2' ~ '<= 2 N comorbidities',
    variable %in% 'comor_3_5' ~ '3 <= N comorbidities <= 5',
    variable %in% 'comor_5mayor' ~ '5 < N comorbidities',
    TRUE ~ 'Unknown'))

surgical_procedure <- surgical_procedure %>% arrange(surgical_procedure_cd) %>%  dplyr::select(surgical_procedure_cd,united_surgical_procedure) 



surgical_procedure <- surgical_procedure %>% mutate(
  descrp = case_when(
    surgical_procedure_cd %in% '0' ~ 'Total',
    surgical_procedure_cd %in% '1' ~ 'Partial',
    surgical_procedure_cd %in% '2' ~ 'Pinning',
    surgical_procedure_cd %in% '3' ~ 'Other',
    TRUE ~ "Non surgical procedure"
    
  )
)

socecon_lvl <- socecon_lvl %>% arrange(socecon_lvl_cd) %>%  dplyr::select(socecon_lvl_cd,united_socecon) %>% 
  mutate(variable = paste0('Socioeconomic lvl: ',socecon_lvl_cd),
         value=united_socecon) %>% dplyr::select(variable,value)


descriptive_values_$variable <- as.character(descriptive_values_$variable)

descriptive_values_$variable[descriptive_values_$variable %in% 'country_cd'] <- 'Country cd'
descriptive_values_$variable[descriptive_values_$variable %in% 'median_age'] <- 'Median age'
descriptive_values_$variable[descriptive_values_$variable %in% 'iqr_age'] <- 'IQR age'
descriptive_values_$variable[descriptive_values_$variable %in% 'n_patient'] <- 'N patients (unique)'
descriptive_values_$variable[descriptive_values_$variable %in% 'n_hospitalization_episode_id'] <- 'N hospitalization episode (unique)'
descriptive_values_$variable[descriptive_values_$variable %in% 'n_patient_with_surgery'] <- 'N patient with surgery (unique)'
descriptive_values_$variable[descriptive_values_$variable %in% 'median_time_adm_to_surgery'] <- 'Median time from admission to surgery (days)'
descriptive_values_$variable[descriptive_values_$variable %in% 'IQR_median_time_adm_to_surgery'] <- 'IQR time from admission to surgery (days)'
descriptive_values_$variable[descriptive_values_$variable %in% 'united_men'] <- 'N Men (%)'
descriptive_values_$variable[descriptive_values_$variable %in% 'united_women'] <- 'N Women (%)'


table <- descriptive_values_ %>% gt(rowname_col = 'variable') %>% 
  tab_header(
    title = "Patient-level summary"
  ) %>% 
  tab_stubhead(label = "Variable") %>% 
     cols_label(
      variable = 'Variable',
      value = 'n (%)'
       ) %>% 
  cols_align(
  align =  "center",
  columns = value)

div(style='height:400px; overflow-y: scroll', table)

socecon_lvl %>% gt(rowname_col = 'variable') %>% 
  tab_header(
    title = "Socioeconomic level table input"
  ) %>% 
     cols_label(
      variable = 'Variable',
      value = 'n (%)'
       ) %>% 
  cols_align(
  align =  "center",
  columns = value)


surgical_procedure %>% 
  dplyr::select(descrp, united_surgical_procedure) %>% gt(rowname_col = 'descrp') %>% 
  tab_header(
    title = "Surgical procedure table input"
  ) %>%
     cols_label(
      descrp = 'Surgical procedure type',
      united_surgical_procedure = 'n (%)'
       ) %>% 
  cols_align(
  align =  "center",
  columns = united_surgical_procedure)


anesthesia_type %>% 
  dplyr::select(descrp, united_anesthesia_type) %>% gt(rowname_col = 'descrp') %>% 
  tab_header(
    title = "Anesthesia type table input"
  ) %>%
     cols_label(
      descrp = 'Anesthesia type',
      united_anesthesia_type = 'n (%)'
       ) %>% 
  cols_align(
  align =  "center",
  columns = united_anesthesia_type)




comorbidities_ %>% 
  dplyr::select(descrp, united_comorbidities) %>% gt(rowname_col = 'descrp') %>% 
  tab_header(
    title = "Number of comorbidities table input"
  ) %>%
     cols_label(
      descrp = 'Number of comorbidities',
      united_comorbidities = 'n (%)'
       ) %>% 
  cols_align(
  align =  "center",
  columns = united_comorbidities)



rm(anesthesia_type,comorbidities_,socecon_lvl,surgical_procedure,descriptive_values_,table)


```


### Table 2. Hospital-level summary

```{r, echo=FALSE,warning=FALSE}
#| label: descriptive tables inputs2


table <- variables_in_hospital %>% gt() %>%
  tab_header(
    title = " Hospital-level summary"
  ) %>%  
  tab_options(
    column_labels.font.size = 8,
  ) %>% 
     cols_label(
      hospital_st = 'Hospital',
      n_patient = 'N patients (unique)',
      n_hospitalization_episode_id = 'N hospitalization episode (unique)',
      n_patient_with_surgery = 'N patient with surgery (unique)',
      median_age = 'Median age',
      iqr_age = 'IQR age',
      women = 'N Women',
      median_time_adm_to_surgery = 'Median time from admission to surgery (days)',
      IQR_median_time_adm_to_surgery = 'IQR time from admission to surgery (days)',
      n_exitus = 'N exitus (all)',
      n_unplanned_readmission_30d = 'N unplanned readmission 30 days (unique)'
       ) %>%
        cols_align(
  align =  "center",
  columns = everything()) %>%
      tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups()) %>%
   opt_interactive(use_filters = TRUE,
                   use_compact_mode = TRUE,
                   use_resizers = TRUE)

write.table(variables_in_hospital,paste0('../../outputs/table_descriptive_hospital_with_unique_patients',params$state,'.csv'),row.names = FALSE, sep='|', na='',fileEncoding="UTF-8")
table
rm(table)
```



## Hospital performance


```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: analysis start

con = dbConnect(duckdb::duckdb(), dbdir=params$data_path, read_only=FALSE)

df_patient <- dbGetQuery(conn = con, "SELECT *,
                         CASE
		WHEN age_nm >= 0
		and age_nm <= 4 THEN 1
		WHEN age_nm >= 5
		and age_nm <= 9 THEN 2
		WHEN age_nm >= 10
		and age_nm <= 14 THEN 3
		WHEN age_nm >= 15
		and age_nm <= 19 THEN 4
		WHEN age_nm >= 20
		and age_nm <= 24 THEN 5
		WHEN age_nm >= 25
		and age_nm <= 29 THEN 6
		WHEN age_nm >= 30
		and age_nm <= 34 THEN 7
		WHEN age_nm >= 35
		and age_nm <= 39 THEN 8
		WHEN age_nm >= 40
		and age_nm <= 44 THEN 9
		WHEN age_nm >= 45
		and age_nm <= 49 THEN 10
		WHEN age_nm >= 50
		and age_nm <= 54 THEN 11
		WHEN age_nm >= 55
		and age_nm <= 59 THEN 12
		WHEN age_nm >= 60
		and age_nm <= 64 THEN 13
		WHEN age_nm >= 65
		and age_nm <= 69 THEN 14
		WHEN age_nm >= 70
		and age_nm <= 74 THEN 15
		WHEN age_nm >= 75
		and age_nm <= 79 THEN 16
		WHEN age_nm >= 80
		and age_nm <= 84 THEN 17
		WHEN age_nm >= 85 
		and age_nm <= 89 THEN 18
		WHEN age_nm >= 90 THEN 19
		ELSE NULL 
		END as age_group,
		FROM patient_cohort")
# df_outpatient_prescription <- dbGetQuery(conn = con, "SELECT * FROM outpatient_prescription")
dbDisconnect(con, shutdown=TRUE)

df_patient <- df_patient %>% mutate_if(is.logical,as.numeric)


start_study <- as.Date('2017-01-01')
end_study <- as.Date('2021-12-31') 

```



```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: analysis continue

df_patient_filter <- df_patient %>% filter(surgical_procedure_bl %in% 1)

df_patient_filter <- df_patient_filter %>% 
  mutate(
    futime_adm_surg = as.numeric(surgical_procedure_dt - hospital_admission_dt),
    futime_surg_end = case_when(
      !is.na(exitus_dt) ~ as.numeric(exitus_dt - surgical_procedure_dt),
      is.na(exitus_dt) ~ as.numeric(end_study - surgical_procedure_dt)),
    exitus_7days_bl = case_when(
      !is.na(exitus_dt) & futime_surg_end<= 7 ~ 1,
      TRUE ~ 0),
      exitus_30days_bl = case_when(
      !is.na(exitus_dt) & futime_surg_end<= 30 ~ 1,
      TRUE ~ 0),
    exitus_hosp_bl = case_when(
      !is.na(exitus_dt) & hospital_discharge_dt>=exitus_dt ~ 1,
      TRUE ~ 0),
    futime_adm_surg_cd = case_when(
      futime_adm_surg <= 1 ~ 1,
      futime_adm_surg > 1 & futime_adm_surg <= 2 ~ 2,
      futime_adm_surg > 2 & futime_adm_surg <= 3 ~ 3,
      TRUE ~ 4)
  )
df_patient_filter <- df_patient_filter %>% filter(futime_adm_surg>=0) # filter remove possible NAs too
df_patient_filter <- df_patient_filter %>% mutate(
  year = year(surgical_procedure_dt),
  month = month(surgical_procedure_dt),
  weekday = weekdays(as.POSIXlt(surgical_procedure_dt),abbreviate = FALSE),
  day_month = day(surgical_procedure_dt)
)


df_model <- df_patient_filter %>% dplyr::select(exitus_hosp_bl,exitus_7days_bl,exitus_30days_bl,unplanned_readmission_30d_bl,unplanned_readmission_7d_bl,futime_adm_surg,futime_adm_surg_cd,age_nm,sex_cd,hospital_st,socecon_lvl_cd,surgical_procedure_cd,anesthesia_type_cd,icu_admission_bl,anticoagulant_status_bl,antiaggregant_status_bl,chronic_kidney_disease_bl,tobacco_copd_bl,depression_bl,serious_mental_illness_bl,alcohol_abuse_bl,             
                                              obesity_overweight_bl,osteoporosis_osteopenia_bl,dementia_bl,diabetes_bl,
                                              liver_disease_bl,pancreatic_disease_bl,inflamatory_bowel_disease_bl,rheumatology_disease_bl,
                                              spinal_cord_disease_injury_bl,serious_neurological_disease_bl,
                                            parkinson_huntington_diseases_bl,seizure_disorder_convulsions_bl,congestive_heart_failure_bl,
                                              coronary_artery_disease_bl,cerebrovascular_disease_bl,peripheral_vascular_disease_bl,
                      traumatic_brain_injury_bl,amputee_bl,year,month,weekday,age_group,hospital_admission_type_cd,previous_hospital_admission_bl)

df_model$sex_cd <- factor(df_model$sex_cd,
levels = c(0,1,2,9),
labels = c("unknown", "male", "female",'not_applicable')) 
df_model$surgical_procedure_cd <- as.integer(df_model$surgical_procedure_cd)
df_model$surgical_procedure_cd <- factor(df_model$surgical_procedure_cd,
                                         levels = c(0,1,2,3),
labels = c("Total", "Partial", "Pinning",'Other')) 
df_model$hospital_st <- as.factor(df_model$hospital_st)
df_model$age_group <- as.factor(df_model$age_group)
df_model$futime_adm_surg_cd <- factor(df_model$futime_adm_surg_cd,
                                         levels = c(1,2,3,4),labels = c("x <=24h", "24h-48h", "48h-72h",">=72h"))
df_model$year <- as.factor(df_model$year)
df_model$month <- as.factor(df_model$month)
#df_model$day_month <- as.factor(df_model$day_month)
### select variables for model with 20% or less of missings
miss_threshold = 0.2
df_model_aux <- df_model %>% select(!c(age_nm,exitus_hosp_bl,exitus_7days_bl,exitus_30days_bl,futime_adm_surg,futime_adm_surg_cd,hospital_st,unplanned_readmission_7d_bl,unplanned_readmission_30d_bl))
a <- names(df_model_aux)[sapply(df_model_aux, function(x) (sum(is.na(x))/nrow(df_model_aux)) <= miss_threshold)]
print(df_model_aux)
df_model_aux <- df_model_aux %>% select(a)
formula_model <- paste0(colnames(df_model_aux), collapse = ' + ')
rm(df_model_aux)
df_model <- df_model %>% select(c(a,age_nm,exitus_hosp_bl,exitus_7days_bl,exitus_30days_bl,futime_adm_surg,futime_adm_surg_cd,hospital_st,unplanned_readmission_7d_bl,unplanned_readmission_30d_bl))

```


### Models

#### Adjusted mortality by time to surgery


##### Adjusted mortality by time to surgery (univariate models)


###### Generalized Linear Model (GLM)


::: {.panel-tabset}

###### Model with variable to predict: exitus in hospital
```{r, echo=FALSE,warning=FALSE}
#| label: GLM exitus in hospital simple

###### exitus in hospital #####


model_exitus_hosp_bl_simple <- glm(exitus_hosp_bl ~ futime_adm_surg_cd, data = df_model, family = binomial(link = "logit"), method = "glm.fit")

summary(model_exitus_hosp_bl_simple)


```


###### Model with variable to predict: exitus in 7 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: GLM exitus 7 days simple

###### exitus 7 days #####

model_exitus_7days_bl_simple <- glm(exitus_7days_bl ~ futime_adm_surg_cd, data = df_model, family = binomial(link = "logit"), method = "glm.fit")

summary(model_exitus_7days_bl_simple)

```


###### Model with variable to predict: exitus in 30 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: GLM exitus 30 days simple

###### exitus 30 days #####


model_exitus_30days_bl_simple <- glm(exitus_30days_bl ~ futime_adm_surg_cd, data = df_model, family = binomial(link = "logit"), method = "glm.fit")

summary(model_exitus_30days_bl_simple)

```

:::

```{r, echo=FALSE,warning=FALSE,output = FALSE}
#| label: filter_hospital

### Filter hospitals by patients

filter_hospital_by_npatients <- function(n_filter, variables_in_hospital, df_model,
                                         n_min_hospital,n_max_hospital){
  variables_in_hospital_filter <- variables_in_hospital %>% filter(n_patient >= n_filter)
  if(length(variables_in_hospital_filter$hospital_st) < n_min_hospital){
    n_total_patients <- sum(variables_in_hospital$n_patient,na.rm=TRUE)
    hospitals_filter <- variables_in_hospital %>% dplyr::select(hospital_st,n_hospitalization_episode_id,n_patient) 
    hospitals_filter <- hospitals_filter[order(hospitals_filter$n_patient,decreasing = TRUE),]
    hospitals_filter <- hospitals_filter %>% mutate(perc_cumsum_patient = 100*(cumsum(n_patient)/n_total_patients))
    #hospitals_filter <- hospitals_filter %>% filter(perc_cumsum_patient <= perc_sample)
    if(nrow(hospitals_filter) > n_max_hospital){
      hospitals_filter <- hospitals_filter[1:n_max_hospital,]
    }
    df_model_filter <- df_model %>% filter(hospital_st %in% hospitals_filter$hospital_st)
  }else{
    if(nrow(variables_in_hospital_filter) > n_max_hospital){
      variables_in_hospital_filter <- variables_in_hospital_filter[order(variables_in_hospital_filter$n_patient,decreasing = TRUE),]
      variables_in_hospital_filter <- variables_in_hospital_filter[1:n_max_hospital,]
    }
    df_model_filter <- df_model %>% filter(hospital_st %in% variables_in_hospital_filter$hospital_st)
  }
  return(df_model_filter)
}
df_model_filter <- filter_hospital_by_npatients(n_filter=50,variables_in_hospital,df_model,n_min_hospital=5, 
                                                n_max_hospital=150)
df_model_filter <- na.omit(df_model_filter)

log_info(paste0("For multilevel model, the number of records is: ",nrow(df_model_filter),' which is a percentage of the sample of: '
,round(100*(nrow(df_model_filter)/sum(variables_in_hospital$n_hospitalization_episode_id, na.rm = TRUE)),2),'%. These models include: ',
length(unique(df_model_filter$hospital_st)),' hospitals of: ',length(unique(variables_in_hospital$hospital_st)),' (',
round(100*(length(unique(df_model_filter$hospital_st))/length(unique(variables_in_hospital$hospital_st))),2),'%)'))


```


###### Generalized Linear Mixed Model (GLMM)

::: {.panel-tabset}



###### Model with variable to predict: exitus in hospital
```{r, echo=FALSE,warning=FALSE}
#| label: GLMM exitus in hospital simple

###### exitus in hospital #####


model_exitus_hosp_bl_simple <- glmer(exitus_hosp_bl ~ futime_adm_surg_cd + (1|hospital_st), data = df_model_filter, family = binomial(link="logit"))

summary(model_exitus_hosp_bl_simple)

df_var <- calculate_ICC(model_exitus_hosp_bl_simple, variable_random_effect = 'hospital_st')


MOR <- round(exp(sqrt(2*(df_var$vcov))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR))


MOR_w_CI <- calculate_MOR(model_exitus_hosp_bl_simple,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI[2],5),' with lower_ci95=',round(MOR_w_CI[1],5),' and upper_ci95=',round(MOR_w_CI[3],5)))



```


###### Model with variable to predict: exitus in 7 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: GLMM exitus 7 days simple

###### exitus 7 days #####

model_exitus_7days_bl_simple <- glmer(exitus_7days_bl ~ futime_adm_surg_cd + (1|hospital_st), data = df_model_filter, family = binomial(link="logit"))

summary(model_exitus_7days_bl_simple)


df_var <- calculate_ICC(model_exitus_7days_bl_simple, variable_random_effect = 'hospital_st')

MOR <- round(exp(sqrt(2*(df_var$vcov))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR))


MOR_w_CI <- calculate_MOR(model_exitus_7days_bl_simple,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI[2],5),' with lower_ci95=',round(MOR_w_CI[1],5),' and upper_ci95=',round(MOR_w_CI[3],5)))



```


###### Model with variable to predict: exitus in 30 days after surgery

```{r, echo=FALSE,warning=FALSE}
#| label: GLMM exitus 30 days simple

###### exitus 30 days #####

model_exitus_30days_bl_simple <- glmer(exitus_30days_bl ~ futime_adm_surg_cd + (1|hospital_st), data = df_model_filter, family = binomial(link="logit"))

summary(model_exitus_30days_bl_simple)

df_var <- calculate_ICC(model_exitus_30days_bl_simple, variable_random_effect = 'hospital_st')

MOR <- round(exp(sqrt(2*(df_var$vcov))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR))



MOR_w_CI <- calculate_MOR(model_exitus_30days_bl_simple,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI[2],5),' with lower_ci95=',round(MOR_w_CI[1],5),' and upper_ci95=',round(MOR_w_CI[3],5)))




```


:::

###### Generalized Additive Model (GAM)

::: {.panel-tabset}

###### Model with variable to predict: exitus in hospital

```{r, echo=FALSE,warning=FALSE}
#| label: GAM exitus in hospital simple

###### exitus in hospital #####


model_exitus_hosp_bl_simple <- mgcv::gam(exitus_hosp_bl ~ s(futime_adm_surg, bs = 'tp'), data = df_model, family = binomial(link="logit"), method = "REML")

summary(model_exitus_hosp_bl_simple)

gratia::draw(model_exitus_hosp_bl_simple,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))
variance_comp(model_exitus_hosp_bl_simple)
```

###### Model with variable to predict: exitus in 7 days after surgery

```{r, echo=FALSE,warning=FALSE}
#| label: GAM exitus 7 days simple

###### exitus 7 days #####

model_exitus_7days_bl_simple <- mgcv::gam(exitus_7days_bl ~ s(futime_adm_surg, bs = 'tp'), data = df_model,family = binomial(link="logit"), method = "REML")

summary(model_exitus_7days_bl_simple)


gratia::draw(model_exitus_7days_bl_simple,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))
variance_comp(model_exitus_7days_bl_simple)
```



###### Model with variable to predict: exitus in 30 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: GAM exitus 30 days simple 

###### exitus 30 days #####


model_exitus_30days_bl_simple <- mgcv::gam(exitus_30days_bl ~ s(futime_adm_surg, bs = 'tp'), data = df_model,family = binomial(link="logit"), method = "REML")


summary(model_exitus_30days_bl_simple)


gratia::draw(model_exitus_30days_bl_simple,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))
variance_comp(model_exitus_30days_bl_simple)

```


:::



###### Generalized Additive Mixed Model (GAMM)


::: {.panel-tabset}

###### Model with variable to predict: exitus in hospital

```{r, echo=FALSE,warning=FALSE}
#| label: GAMM exitus in hospital simple

###### exitus in hospital #####

model_exitus_hosp_bl_simple <- mgcv::gam(exitus_hosp_bl ~ s(futime_adm_surg, bs = 'tp') + s(hospital_st, bs = 're'), data = df_model_filter,family=binomial(link="logit"), method = "REML")

summary(model_exitus_hosp_bl_simple)



gratia::draw(model_exitus_hosp_bl_simple,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))


df_var_re <- calculate_ICC(model_exitus_hosp_bl_simple, variable_random_effect = 'hospital_st')



MOR <- round(exp(sqrt(2*(df_var_re$variance))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR))


MOR_w_CI <- calculate_MOR(model_exitus_hosp_bl_simple,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI[2],5),' with lower_ci95=',round(MOR_w_CI[1],5),' and upper_ci95=',round(MOR_w_CI[3],5)))


```

###### Model with variable to predict: exitus in 7 days after surgery

```{r, echo=FALSE,warning=FALSE}
#| label: GAMM exitus 7 days simple

###### exitus 7 days #####


model_exitus_7days_bl_simple <- mgcv::gam(exitus_7days_bl ~ s(futime_adm_surg, bs = 'tp') + s(hospital_st, bs = 're'), data = df_model_filter,family=binomial(link="logit"), method = "REML")

summary(model_exitus_7days_bl_simple)



gratia::draw(model_exitus_7days_bl_simple,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))


df_var_re <- calculate_ICC(model_exitus_7days_bl_simple, variable_random_effect = 'hospital_st')

MOR <- round(exp(sqrt(2*(df_var_re$variance))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR))


MOR_w_CI <- calculate_MOR(model_exitus_7days_bl_simple,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI[2],5),' with lower_ci95=',round(MOR_w_CI[1],5),' and upper_ci95=',round(MOR_w_CI[3],5)))




```



###### Model with variable to predict: exitus in 30 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: GAMM exitus 30 days simple

###### exitus 30 days #####


model_exitus_30days_bl_simple <- mgcv::gam(exitus_30days_bl ~ s(futime_adm_surg, bs = 'tp') + s(hospital_st, bs = 're'), data = df_model_filter,family=binomial(link="logit"), method = "REML")


summary(model_exitus_30days_bl_simple)



gratia::draw(model_exitus_30days_bl_simple,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))



df_var_re <- calculate_ICC(model_exitus_30days_bl_simple, variable_random_effect = 'hospital_st')

MOR <- round(exp(sqrt(2*(df_var_re$variance))*qnorm(.75)),5)

print(paste0('Median odds ratio (MOR): ', MOR))


MOR_w_CI <- calculate_MOR(model_exitus_30days_bl_simple,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI[2],5),' with lower_ci95=',round(MOR_w_CI[1],5),' and upper_ci95=',round(MOR_w_CI[3],5)))



```


:::



##### Adjusted mortality by time to surgery (multivariate models)


::: {.panel-tabset}

###### Model with variable to predict: exitus in hospital


```{r, echo=FALSE,warning=FALSE}
#| label: GAMM exitus in hospital complex model

###### exitus in hospital #####

formula_model_exitus_hosp_bl <- as.formula(paste0('exitus_hosp_bl ~ s(futime_adm_surg, bs = "tp") + ',formula_model, ' + s(hospital_st, bs = "re")'))


model_exitus_hosp_bl <- mgcv::gam(formula_model_exitus_hosp_bl, data = df_model_filter,family=binomial(link="logit"), method = "REML")

summary(model_exitus_hosp_bl)



gratia::draw(model_exitus_hosp_bl,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))

##### Not use function 'calculate_ICC' here to save variables #####

var <- variance_comp(model_exitus_hosp_bl)
var
var_re_exitus_inhospital <- var %>% filter(component %in% "s(hospital_st)")

icc_exitus_inhospital <- (var_re_exitus_inhospital$variance/(var_re_exitus_inhospital$variance + ((pi*pi)/3)))
if(sum(is.na(icc_exitus_inhospital))== 0){
  
icc_lower_exitus_inhospital <- (var_re_exitus_inhospital$variance - var_re_exitus_inhospital$std_dev)/((var_re_exitus_inhospital$variance - var_re_exitus_inhospital$std_dev)+((pi*pi)/3))
icc_upper_exitus_inhospital <- (var_re_exitus_inhospital$variance + var_re_exitus_inhospital$std_dev)/((var_re_exitus_inhospital$variance + var_re_exitus_inhospital$std_dev)+((pi*pi)/3))
             
print(paste0('ICC ajusted: ', round(icc_exitus_inhospital,5),' with lower_ci95=',round(icc_lower_exitus_inhospital,5),' and upper_ci95=',round(icc_upper_exitus_inhospital,5)))  
}else{
  print(paste0('ICC ajusted: ', 'It has not been possible to calculate'))
}

MOR_exitus_inhospital <- round(exp(sqrt(2*(var_re_exitus_inhospital$variance))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR_exitus_inhospital))


MOR_w_CI_exitus_inhospital <- calculate_MOR(model_exitus_hosp_bl,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI_exitus_inhospital[2],5),
             ' with lower_ci95=',round(MOR_w_CI_exitus_inhospital[1],5),' and upper_ci95=',round(MOR_w_CI_exitus_inhospital[3],5)))




df_model_filter[c('predict_model_exitus_hospital_bl','se_predict_model_exitus_hospital_bl')] <- model_exitus_hosp_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)


df_model_filter$resid_model_exitus_hospital_bl <- resid(model_exitus_hosp_bl,type='response')

output_model_exitus_hospital_bl <- df_model_filter

df_patient_ <- df_patient %>% filter(!is.na(country_cd))
output_model_exitus_hospital_bl$country_cd <- unique(df_patient_$country_cd)
output_model_exitus_hospital_bl <- output_model_exitus_hospital_bl %>% 
  group_by(country_cd,hospital_st) %>% 
  summarise(
            mean_predict_model = mean(predict_model_exitus_hospital_bl,na.rm=TRUE),
            sd_predict_model= sd(predict_model_exitus_hospital_bl,na.rm=TRUE),
            mean_residuals_model = mean(resid_model_exitus_hospital_bl,na.rm=TRUE),
            sd_residuals_model = sd(resid_model_exitus_hospital_bl,na.rm=TRUE),
            median_predict_model = median(predict_model_exitus_hospital_bl,na.rm=TRUE),
            P25_predict_model = round(quantile(predict_model_exitus_hospital_bl,0.25,na.rm=TRUE),5),
            P75_predict_model = round(quantile(predict_model_exitus_hospital_bl,0.75,na.rm=TRUE),5),
            P05_predict_model = round(quantile(predict_model_exitus_hospital_bl,0.05,na.rm=TRUE),5),
            P95_predict_model = round(quantile(predict_model_exitus_hospital_bl,0.95,na.rm=TRUE),5),
            median_residuals_model = median(resid_model_exitus_hospital_bl,na.rm=TRUE),
            P25_residuals_model = round(quantile(resid_model_exitus_hospital_bl,0.25,na.rm=TRUE),5),
            P75_residuals_model = round(quantile(resid_model_exitus_hospital_bl,0.75,na.rm=TRUE),5),
            P05_residuals_model = round(quantile(resid_model_exitus_hospital_bl,0.05,na.rm=TRUE),5),
            P95_residuals_model = round(quantile(resid_model_exitus_hospital_bl,0.95,na.rm=TRUE),5)
            )

write.table(output_model_exitus_hospital_bl,paste0('../../outputs/aggregated_outputs_byhospital_GAMM_exitus_inhospital',params$state,'.csv'), sep='|',row.names=FALSE,na='',fileEncoding="UTF-8")

ggplot(df_model_filter, aes(x = resid_model_exitus_hospital_bl, y = hospital_st, group = hospital_st)) + 
  geom_density_ridges(alpha=0.7) +
  labs(title = 'Density plot of residuals for each hospital',y= 'Hospital',x='Residuals') +
  theme(panel.background = element_blank(),
        axis.line = element_line(color='black'))
rm(output_model_exitus_hospital_bl)
```



###### Model with variable to predict: exitus in 7 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: GAMM exitus 7 days complex model

###### exitus in 7 days #####

formula_model_exitus_7days_bl <- as.formula(paste0('exitus_7days_bl ~ s(futime_adm_surg, bs = "tp") + ',formula_model, ' + s(hospital_st, bs = "re")'))

model_exitus_7days_bl <- mgcv::gam(formula_model_exitus_7days_bl, data = df_model_filter,family=binomial(link="logit"), method = "REML")


summary(model_exitus_7days_bl)



gratia::draw(model_exitus_7days_bl,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))

##### Not use function 'calculate_ICC' here to save variables #####

var <- variance_comp(model_exitus_7days_bl)
var
var_re_exitus_7d <- var %>% filter(component %in% "s(hospital_st)")

icc_exitus_7d <- (var_re_exitus_7d$variance/(var_re_exitus_7d$variance + ((pi*pi)/3)))
if(sum(is.na(icc_exitus_7d))== 0){
  
icc_lower_exitus_7d <- (var_re_exitus_7d$variance - var_re_exitus_7d$std_dev)/((var_re_exitus_7d$variance - var_re_exitus_7d$std_dev)+((pi*pi)/3))
icc_upper_exitus_7d <- (var_re_exitus_7d$variance + var_re_exitus_7d$std_dev)/((var_re_exitus_7d$variance + var_re_exitus_7d$std_dev)+((pi*pi)/3))
             
print(paste0('ICC ajusted: ', round(icc_exitus_7d,5),' with lower_ci95=',round(icc_lower_exitus_7d,5),' and upper_ci95=',round(icc_upper_exitus_7d,5)))  
}else{
  print(paste0('ICC ajusted: ', 'It has not been possible to calculate'))
}

MOR_exitus_7d <- round(exp(sqrt(2*(var_re_exitus_7d$variance))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR_exitus_7d))




MOR_w_CI_exitus_7d <- calculate_MOR(model_exitus_7days_bl,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI_exitus_7d[2],5),
             ' with lower_ci95=',round(MOR_w_CI_exitus_7d[1],5),' and upper_ci95=',round(MOR_w_CI_exitus_7d[3],5)))




df_model_filter[c('predict_model_exitus_7days_bl','se_predict_model_exitus_7days_bl')] <- model_exitus_7days_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)

df_model_filter$resid_model_exitus_7days_bl <- resid(model_exitus_7days_bl,type='response')

output_model_exitus_7days_bl <- df_model_filter

df_patient_ <- df_patient %>% filter(!is.na(country_cd))
output_model_exitus_7days_bl$country_cd <- unique(df_patient_$country_cd)

output_model_exitus_7days_bl <- output_model_exitus_7days_bl %>% 
  group_by(country_cd,hospital_st) %>% 
  summarise(
            mean_predict_model = mean(predict_model_exitus_7days_bl,na.rm=TRUE),
            sd_predict_model = sd(predict_model_exitus_7days_bl,na.rm=TRUE),
            mean_residuals_model = mean(resid_model_exitus_7days_bl,na.rm=TRUE),
            sd_residuals_model = sd(resid_model_exitus_7days_bl,na.rm=TRUE),
            median_predict_model = median(predict_model_exitus_7days_bl,na.rm=TRUE),
            P25_predict_model = round(quantile(predict_model_exitus_7days_bl,0.25,na.rm=TRUE),5),
            P75_predict_model = round(quantile(predict_model_exitus_7days_bl,0.75,na.rm=TRUE),5),
            P05_predict_model = round(quantile(predict_model_exitus_7days_bl,0.05,na.rm=TRUE),5),
            P95_predict_model = round(quantile(predict_model_exitus_7days_bl,0.95,na.rm=TRUE),5),
            median_residuals_model = median(resid_model_exitus_7days_bl,na.rm=TRUE),
            P25_residuals_model = round(quantile(resid_model_exitus_7days_bl,0.25,na.rm=TRUE),5),
            P75_residuals_model = round(quantile(resid_model_exitus_7days_bl,0.75,na.rm=TRUE),5),
            P05_residuals_model = round(quantile(resid_model_exitus_7days_bl,0.05,na.rm=TRUE),5),
            P95_residuals_model = round(quantile(resid_model_exitus_7days_bl,0.95,na.rm=TRUE),5)
            )

write.table(output_model_exitus_7days_bl,paste0('../../outputs/aggregated_outputs_byhospital_GAMM_exitus_in7days',params$state,'.csv'), sep='|',row.names=FALSE,na='',fileEncoding="UTF-8")

ggplot(df_model_filter, aes(x = resid_model_exitus_7days_bl, y = hospital_st, group = hospital_st)) + 
  geom_density_ridges(alpha=0.7) +
  labs(title = 'Density plot of residuals for each hospital',y= 'Hospital',x='Residuals') +
  theme(panel.background = element_blank(),
        axis.line = element_line(color='black'))

rm(output_model_exitus_7days_bl)
```

###### Model with variable to predict: exitus in 30 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: GAMM exitus 30 days complex 

###### exitus 30 days #####

formula_model_exitus_30days_bl <- as.formula(paste0('exitus_30days_bl ~ s(futime_adm_surg, bs = "tp") + ',formula_model, ' + s(hospital_st, bs = "re")'))

model_exitus_30days_bl <- mgcv::gam(formula_model_exitus_30days_bl, data = df_model_filter,family=binomial(link="logit"), method = "REML")

summary(model_exitus_30days_bl)



gratia::draw(model_exitus_30days_bl,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))

##### Not use function 'calculate_ICC' here to save variables #####

var <- variance_comp(model_exitus_30days_bl)
var
var_re_exitus_30d <- var %>% filter(component %in% "s(hospital_st)")

icc_exitus_30d <- (var_re_exitus_30d$variance/(var_re_exitus_30d$variance + ((pi*pi)/3)))
if(sum(is.na(icc_exitus_30d))== 0){
  
icc_lower_exitus_30d <- (var_re_exitus_30d$variance - var_re_exitus_30d$std_dev)/((var_re_exitus_30d$variance - var_re_exitus_30d$std_dev)+((pi*pi)/3))
icc_upper_exitus_30d <- (var_re_exitus_30d$variance + var_re_exitus_30d$std_dev)/((var_re_exitus_30d$variance + var_re_exitus_30d$std_dev)+((pi*pi)/3))
             
print(paste0('ICC ajusted: ', round(icc_exitus_30d,5),' with lower_ci95=',round(icc_lower_exitus_30d,5),' and upper_ci95=',round(icc_upper_exitus_30d,5)))   

}else{
  print(paste0('ICC ajusted: ', 'It has not been possible to calculate'))
}

MOR_exitus_30d <- round(exp(sqrt(2*(var_re_exitus_30d$variance))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR_exitus_30d))



MOR_w_CI_exitus_30d <- calculate_MOR(model_exitus_30days_bl,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI_exitus_30d[2],5),
             ' with lower_ci95=',round(MOR_w_CI_exitus_30d[1],5),' and upper_ci95=',round(MOR_w_CI_exitus_30d[3],5)))







df_model_filter[c('predict_model_exitus_30days_bl','se_predict_model_exitus_30days_bl')] <- model_exitus_30days_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)


df_model_filter$resid_model_exitus_30days_bl <- resid(model_exitus_30days_bl,type='response')


output_model_exitus_30days_bl <- df_model_filter

df_patient_ <- df_patient %>% filter(!is.na(country_cd))
output_model_exitus_30days_bl$country_cd <- unique(df_patient_$country_cd)


output_model_exitus_30days_bl <- output_model_exitus_30days_bl %>% 
  group_by(country_cd,hospital_st) %>% 
  summarise(
            mean_predict_model = mean(predict_model_exitus_30days_bl,na.rm=TRUE),
            sd_predict_model = sd(predict_model_exitus_30days_bl,na.rm=TRUE),
            mean_residuals_model = mean(resid_model_exitus_30days_bl,na.rm=TRUE),
            sd_residuals_model = sd(resid_model_exitus_30days_bl,na.rm=TRUE),
            median_predict_model = median(predict_model_exitus_30days_bl,na.rm=TRUE),
            P25_predict_model = round(quantile(predict_model_exitus_30days_bl,0.25,na.rm=TRUE),5),
            P75_predict_model = round(quantile(predict_model_exitus_30days_bl,0.75,na.rm=TRUE),5),
            P05_predict_model = round(quantile(predict_model_exitus_30days_bl,0.05,na.rm=TRUE),5),
            P95_predict_model = round(quantile(predict_model_exitus_30days_bl,0.95,na.rm=TRUE),5),
            median_residuals_model = median(resid_model_exitus_30days_bl,na.rm=TRUE),
            P25_residuals_model = round(quantile(resid_model_exitus_30days_bl,0.25,na.rm=TRUE),5),
            P75_residuals_model = round(quantile(resid_model_exitus_30days_bl,0.75,na.rm=TRUE),5),
            P05_residuals_model = round(quantile(resid_model_exitus_30days_bl,0.05,na.rm=TRUE),5),
            P95_residuals_model = round(quantile(resid_model_exitus_30days_bl,0.95,na.rm=TRUE),5)
            )

write.table(output_model_exitus_30days_bl,paste0('../../outputs/aggregated_outputs_byhospital_GAMM_exitus_in30days',params$state,'.csv'), sep='|',row.names=FALSE,na='',fileEncoding="UTF-8")

ggplot(df_model_filter, aes(x = resid_model_exitus_30days_bl, y = hospital_st, group = hospital_st)) + 
  geom_density_ridges(alpha=0.7) +
  labs(title = 'Density plot of residuals for each hospital',y= 'Hospital',x='Residuals') +
  theme(panel.background = element_blank(),
        axis.line = element_line(color='black'))
rm(output_model_exitus_30days_bl)
```


:::


###### Performance comparison of multivariate GAMM models 


```{r, warning=FALSE, echo=FALSE}
#| label: performance model exitus

p <- compare_performance(model_exitus_hosp_bl,model_exitus_7days_bl, model_exitus_30days_bl)

p$Name[p$Name %in% 'model_exitus_hosp_bl'] <- 'Model 1: exitus in hospital'
p$Name[p$Name %in% 'model_exitus_7days_bl'] <- 'Model 2: exitus in 7 days after surgery'
p$Name[p$Name %in% 'model_exitus_30days_bl'] <- 'Model 3: exitus in 30 days after surgery'

p
write.table(p,paste0('../../outputs/performance_GAMM_exitus',params$state,'.csv'), sep='|',row.names=FALSE,na='',fileEncoding="UTF-8")
```
 

#### Adjusted unplanned readmissions by time to surgery (multivariate models)


::: {.panel-tabset}

##### Model with variable to predict: unplanned readmission (for any cause) in the next 7 days after hospitalization

```{r, echo=FALSE,warning=FALSE}
#| label: GAMM unplanned 7 days complex model

##### unplanned 7 days ######

formula_model_unplanned_7days_bl <- as.formula(paste0('unplanned_readmission_7d_bl ~ s(futime_adm_surg, bs = "tp") + ',formula_model, ' + s(hospital_st, bs = "re")'))

model_unplanned_7days_bl <- mgcv::gam(formula_model_unplanned_7days_bl, data = df_model_filter,family=binomial(link="logit"), method = "REML")

summary(model_unplanned_7days_bl)


gratia::draw(model_unplanned_7days_bl,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))

##### Not use function 'calculate_ICC' here to save variables #####

var <- variance_comp(model_unplanned_7days_bl)
var
var_re_unplanned_7d <- var %>% filter(component %in% "s(hospital_st)")

icc_unplanned_7d <- (var_re_unplanned_7d$variance/(var_re_unplanned_7d$variance + ((pi*pi)/3)))
if(sum(is.na(icc_unplanned_7d))== 0){
  
icc_lower_unplanned_7d <- (var_re_unplanned_7d$variance - var_re_unplanned_7d$std_dev)/((var_re_unplanned_7d$variance - var_re_unplanned_7d$std_dev)+((pi*pi)/3))
icc_upper_unplanned_7d <- (var_re_unplanned_7d$variance + var_re_unplanned_7d$std_dev)/((var_re_unplanned_7d$variance + var_re_unplanned_7d$std_dev)+((pi*pi)/3))
             
print(paste0('ICC ajusted: ', round(icc_unplanned_7d,5),' with lower_ci95=',round(icc_lower_unplanned_7d,5),' and upper_ci95=',round(icc_upper_unplanned_7d,5)))    

}else{
  print(paste0('ICC ajusted: ', 'It has not been possible to calculate'))
}

MOR_unplanned_7d <- round(exp(sqrt(2*(var_re_unplanned_7d$variance))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR_unplanned_7d))



MOR_w_CI_unplanned_7d <- calculate_MOR(model_unplanned_7days_bl,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI_unplanned_7d[2],5),
             ' with lower_ci95=',round(MOR_w_CI_unplanned_7d[1],5),' and upper_ci95=',round(MOR_w_CI_unplanned_7d[3],5)))





df_model_filter[c('predict_model_unplanned_7days_bl','se_predict_model_unplanned_7days_bl')] <- model_unplanned_7days_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)

df_model_filter$resid_model_unplanned_7days_bl <- resid(model_unplanned_7days_bl,type='response')


output_model_unplanned_7days_bl <- df_model_filter

df_patient_ <- df_patient %>% filter(!is.na(country_cd))
output_model_unplanned_7days_bl$country_cd <- unique(df_patient_$country_cd)

output_model_unplanned_7days_bl <- output_model_unplanned_7days_bl %>% 
  group_by(country_cd,hospital_st) %>% 
  summarise(
            mean_predict_model = mean(predict_model_unplanned_7days_bl,na.rm=TRUE),
            sd_predict_model = sd(predict_model_unplanned_7days_bl,na.rm=TRUE),
            mean_residuals_model = mean(resid_model_unplanned_7days_bl,na.rm=TRUE),
            sd_residuals_model = sd(resid_model_unplanned_7days_bl,na.rm=TRUE),
            median_predict_model = median(predict_model_unplanned_7days_bl,na.rm=TRUE),
            P25_predict_model = round(quantile(predict_model_unplanned_7days_bl,0.25,na.rm=TRUE),5),
            P75_predict_model = round(quantile(predict_model_unplanned_7days_bl,0.75,na.rm=TRUE),5),
            P05_predict_model = round(quantile(predict_model_unplanned_7days_bl,0.05,na.rm=TRUE),5),
            P95_predict_model = round(quantile(predict_model_unplanned_7days_bl,0.95,na.rm=TRUE),5),
            median_residuals_model = median(resid_model_unplanned_7days_bl,na.rm=TRUE),
            P25_residuals_model = round(quantile(resid_model_unplanned_7days_bl,0.25,na.rm=TRUE),5),
            P75_residuals_model = round(quantile(resid_model_unplanned_7days_bl,0.75,na.rm=TRUE),5),
            P05_residuals_model = round(quantile(resid_model_unplanned_7days_bl,0.05,na.rm=TRUE),5),
            P95_residuals_model = round(quantile(resid_model_unplanned_7days_bl,0.95,na.rm=TRUE),5)
            )


write.table(output_model_unplanned_7days_bl,paste0('../../outputs/aggregated_outputs_byhospital_GAMM_unplanned_7days',params$state,'.csv'), sep='|',row.names=FALSE)

ggplot(df_model_filter, aes(x = resid_model_unplanned_7days_bl, y = hospital_st, group = hospital_st)) + 
  geom_density_ridges(alpha=0.7) +
  labs(title = 'Density plot of residuals for each hospital',y= 'Hospital',x='Residuals') +
  theme(panel.background = element_blank(),
        axis.line = element_line(color='black'))
rm(output_model_unplanned_7days_bl)
```


##### Model with variable to predict: unplanned readmission (for any cause) in the next 30 days after hospitalization


```{r, echo=FALSE,warning=FALSE}
#| label: GAMM unplanned 30 days complex model

###### unplanned 30 days #####

formula_model_unplanned_30days_bl <- as.formula(paste0('unplanned_readmission_30d_bl ~ s(futime_adm_surg, bs = "tp") + ',formula_model, ' + s(hospital_st, bs = "re")'))

model_unplanned_30days_bl <- mgcv::gam(formula_model_unplanned_30days_bl, data = df_model_filter,family=binomial(link="logit"), method = "REML")

summary(model_unplanned_30days_bl)


gratia::draw(model_unplanned_30days_bl,select = 1,ci_level = 0.95) + 
  geom_hline(yintercept = 0,colour = 'red', lty=2) + 
  labs(title = 'Time from admission to surgery (spline)',x='Time from admission to surgery (days)') +
  theme(panel.background = element_blank(), axis.line = element_line(color = 'black'))

##### Not use function 'calculate_ICC' here to save variables #####

var <- variance_comp(model_unplanned_30days_bl)
var
var_re_unplanned_30d <- var %>% filter(component %in% "s(hospital_st)")

icc_unplanned_30d <- (var_re_unplanned_30d$variance/(var_re_unplanned_30d$variance + ((pi*pi)/3)))
if(sum(is.na(icc_unplanned_30d))== 0){
  
  
icc_lower_unplanned_30d <- (var_re_unplanned_30d$variance - var_re_unplanned_30d$std_dev)/((var_re_unplanned_30d$variance - var_re_unplanned_30d$std_dev)+((pi*pi)/3))
icc_upper_unplanned_30d <- (var_re_unplanned_30d$variance + var_re_unplanned_30d$std_dev)/((var_re_unplanned_30d$variance + var_re_unplanned_30d$std_dev)+((pi*pi)/3))
             
print(paste0('ICC ajusted: ', round(icc_unplanned_30d,5),' with lower_ci95=',round(icc_lower_unplanned_30d,5),' and upper_ci95=',round(icc_upper_unplanned_30d,5)))    

}else{
  print(paste0('ICC ajusted: ', 'It has not been possible to calculate'))
}


MOR_unplanned_30d <- round(exp(sqrt(2*(var_re_unplanned_30d$variance))*qnorm(.75)),5)
print(paste0('Median odds ratio (MOR): ', MOR_unplanned_30d))



MOR_w_CI_unplanned_30d <- calculate_MOR(model_unplanned_30days_bl,var_re = 'hospital_st',n_iter = 1000)
print(paste0('MOR with bootsrapping: ',round(MOR_w_CI_unplanned_30d[2],5),
             ' with lower_ci95=',round(MOR_w_CI_unplanned_30d[1],5),' and upper_ci95=',round(MOR_w_CI_unplanned_30d[3],5)))



df_model_filter[c('predict_model_unplanned_30days_bl','se_predict_model_unplanned_30days_bl')] <- model_unplanned_30days_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)

df_model_filter$resid_model_unplanned_30days_bl <- resid(model_unplanned_30days_bl,type='response')


output_model_unplanned_30days_bl <- df_model_filter

df_patient_ <- df_patient %>% filter(!is.na(country_cd))
output_model_unplanned_30days_bl$country_cd <- unique(df_patient_$country_cd)

output_model_unplanned_30days_bl <- output_model_unplanned_30days_bl %>% 
  group_by(country_cd,hospital_st) %>% 
  summarise(
            mean_predict_model = mean(predict_model_unplanned_30days_bl,na.rm=TRUE),
            sd_predict_model = sd(predict_model_unplanned_30days_bl,na.rm=TRUE),
            mean_residuals_model = mean(resid_model_unplanned_30days_bl,na.rm=TRUE),
            sd_residuals_model = sd(resid_model_unplanned_30days_bl,na.rm=TRUE),
            median_predict_model = median(predict_model_unplanned_30days_bl,na.rm=TRUE),
            P25_predict_model = round(quantile(predict_model_unplanned_30days_bl,0.25,na.rm=TRUE),5),
            P75_predict_model = round(quantile(predict_model_unplanned_30days_bl,0.75,na.rm=TRUE),5),
            P05_predict_model = round(quantile(predict_model_unplanned_30days_bl,0.05,na.rm=TRUE),5),
            P95_predict_model = round(quantile(predict_model_unplanned_30days_bl,0.95,na.rm=TRUE),5),
            median_residuals_model = median(resid_model_unplanned_30days_bl,na.rm=TRUE),
            P25_residuals_model = round(quantile(resid_model_unplanned_30days_bl,0.25,na.rm=TRUE),5),
            P75_residuals_model = round(quantile(resid_model_unplanned_30days_bl,0.75,na.rm=TRUE),5),
            P05_residuals_model = round(quantile(resid_model_unplanned_30days_bl,0.05,na.rm=TRUE),5),
            P95_residuals_model = round(quantile(resid_model_unplanned_30days_bl,0.95,na.rm=TRUE),5)
            )

write.table(output_model_unplanned_30days_bl,paste0('../../outputs/aggregated_outputs_byhospital_GAMM_unplanned_30days',params$state,'.csv'), sep='|',row.names=FALSE)

ggplot(df_model_filter, aes(x = resid_model_unplanned_30days_bl, y = hospital_st, group = hospital_st)) + 
  geom_density_ridges(alpha=0.7) +
  labs(title = 'Density plot of residuals for each hospital',y= 'Hospital',x='Residuals') +
  theme(panel.background = element_blank(),
        axis.line = element_line(color='black'))
rm(output_model_unplanned_30days_bl)
```


:::


##### Performance comparison of multivariate GAMM models 


```{r, warning=FALSE, echo=FALSE}
#| label: performance model unplanned

p <- compare_performance(model_unplanned_7days_bl, model_unplanned_30days_bl)


p$Name[p$Name %in% 'model_unplanned_7days_bl'] <- 'Model 1: unplanned readmission in the next 7 days after hospitalization'
p$Name[p$Name %in% 'model_unplanned_30days_bl'] <- 'Model 2: unplanned readmission in the next 30 days after hospitalization'

p
write.table(p,paste0('../../outputs/performance_GAMM_unplanned_readmission',params$state,'.csv'), sep='|',row.names=FALSE,na='',fileEncoding="UTF-8")
```



```{r,echo=FALSE,warning=FALSE,output=FALSE}
#| label: save summary

source('./aux_script/save_summary.R')


```


