---
title: "ICCONIC hip fracture: Analytical pipeline"
author: Data Science for Health Services and Policy Research (IACS)
editor: source

#date: 
# bibliography: 
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 5
    highlight-style: pygments
    code-fold: true
    html-math-method: katex
execute: 
  warning: false
  cache: false
  echo: false
params:
    data_path: '../../inputs/data.duckdb' 
---

```{css, echo = FALSE}
.justify {
  text-align: justify !important
}
```


```{r, output = FALSE, echo=FALSE, warning=FALSE}
#| label: load r packages
library(duckdb)
library(dplyr)
library(Hmisc)
library(survival)
library(survminer)
library(lubridate)
library(mgcv)
library(stringr)
library(sjPlot)
library(gt)
library(reshape2)
library(htmltools)
# library(ggplotify)
```



```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: query descriptive


con = dbConnect(duckdb::duckdb(), dbdir=params$data_path, read_only=FALSE)

descriptive_values <- dbGetQuery(conn = con, "
SELECT median(age_nm) as median_age, (QUANTILE_CONT(age_nm, 0.75)-QUANTILE_CONT(age_nm, 0.25)) as iqr_age,
count(distinct patient_id) filter (where sex_cd = '1') as men,
count(distinct patient_id) filter (where sex_cd = '2') as women,
count(distinct patient_id) as n_patient,
count(distinct hospitalization_episode_id) as n_hospitalization_episode_id,
count(distinct patient_id) filter (where surgical_procedure_bl = TRUE) as n_patient_with_surgery,
median(surgical_procedure_dt -hospital_admission_dt) as median_time_adm_to_surgery,
(QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.75)-QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.25)) as IQR_median_time_adm_to_surgery,
-- round(mean(surgical_procedure_dt -hospital_admission_dt),2) as mean_time_adm_to_surgery
FROM main.patient_cohort")
descriptive_values <- descriptive_values %>% mutate(
  perc_men = round((men/n_patient)*100,2),
  united_men =paste0(men, ' (',perc_men,'%)' ),
  perc_women = round((women/n_patient)*100,2),
  united_women =paste0(women, ' (',perc_women,'%)' )
)
descriptive_values <- descriptive_values %>% select(!c(men,women,perc_men,perc_women))
## socec

socecon_lvl <- dbGetQuery(conn = con,"
SELECT
	coalesce(socecon_lvl_cd, 'Unknown') as socecon_lvl_cd,
	count(distinct patient_id) as patient_distinct_socecon,
	
FROM
	main.patient_cohort
GROUP BY
	socecon_lvl_cd ")
socecon_lvl <- socecon_lvl %>% mutate(
  perc_socecon_lvl = round((patient_distinct_socecon/descriptive_values$n_patient)*100,2),
  united_socecon =paste0(patient_distinct_socecon, ' (',perc_socecon_lvl,'%)' )
)

## surgical_procedure_cd
surgical_procedure <- dbGetQuery(conn = con,"
SELECT
	coalesce(surgical_procedure_cd::int,
	'Non surgical procedure' ) as surgical_procedure_cd ,
	count(distinct patient_id) as patient_distinct_surgical_procedure_cd
FROM
	main.patient_cohort
GROUP BY
	surgical_procedure_cd")
	
surgical_procedure <- surgical_procedure %>% mutate(
  perc_surgical_procedure = round((patient_distinct_surgical_procedure_cd/descriptive_values$n_patient)*100,2),
  united_surgical_procedure=paste0(patient_distinct_surgical_procedure_cd, ' (',perc_surgical_procedure,'%)' )
)

## anesthesia_type_cd only in patients with surgical procedure, so perc is divided by number of patients with surgery

anesthesia_type <- dbGetQuery(conn=con,"
 SELECT
	coalesce(anesthesia_type_cd::int,
	'Unknown') as anesthesia_type_cd,
	count(distinct patient_id) as patient_distinct_anesthesia_type_cd
FROM
	main.patient_cohort
WHERE surgical_procedure_bl = TRUE
GROUP BY
	anesthesia_type_cd ")

anesthesia_type <- anesthesia_type %>% mutate(
  perc_anesthesia_type = round((patient_distinct_anesthesia_type_cd/descriptive_values$n_patient_with_surgery)*100,2),
  united_anesthesia_type=paste0(patient_distinct_anesthesia_type_cd, ' (',perc_anesthesia_type,'%)' )
)


comorbidities <- dbGetQuery(conn=con,"
with seleccion as (
SELECT
	patient_id ,
	list_aggregate([chronic_kidney_disease_bl::int,
	tobacco_copd_bl::int,
	depression_bl::int,
	serious_mental_illness_bl::int,
	alcohol_abuse_bl::int,
	obesity_overweight_bl::int,
	osteoporosis_osteopenia_bl::int,
	dementia_bl::int,
	diabetes_bl::int,
	liver_disease_bl::int,
	pancreatic_disease_bl::int,
	inflamatory_bowel_disease_bl::int,
	rheumatology_disease_bl::int,
	spinal_cord_disease_injury_bl::int,
	serious_neurological_disease_bl::int,
	parkinson_huntintong_diseases_bl::int,
	seizure_disorder_convulsions_bl::int,
	congestive_heart_failure_bl::int,
	coronary_artery_disease_bl::int,
	cerebrovascular_disease_bl::int,
	peripheral_vascular_disease_bl::int,
	traumatic_brain_injury_bl::int,
	amputee_bl::int],
	'sum') as sum_comorbidities
from
	main.patient_cohort)
select
	sum_comorbidities,
	count(DISTINCT patient_id) as patient_distinct_n_comorbidities
from
	seleccion
group by
	sum_comorbidities")

comorbidities <- comorbidities %>% mutate(
  perc_comorbidities = round((patient_distinct_n_comorbidities/descriptive_values$n_patient)*100,2),
  united_comorbidities=paste0(patient_distinct_n_comorbidities, ' (',perc_comorbidities,'%)' )
)


## surgical_procedure_cd in socecon_lvl_cd; percent with denominador n_patient distinct of each socecon level

surgical_procedure_in_socecon_lvl  <- dbGetQuery(conn=con,"
SELECT
	coalesce(socecon_lvl_cd,
	'Unknown') as socecon_lvl_cd,
	coalesce(surgical_procedure_cd::int,
	'Non surgical procedure' ) as surgical_procedure_cd ,
	count(DISTINCT patient_id) as n_patient_surgical_socecon
from
	main.patient_cohort
group by
	socecon_lvl_cd ,
	surgical_procedure_cd")
	
surgical_procedure_in_socecon_lvl <- left_join(x=surgical_procedure_in_socecon_lvl, y=socecon_lvl[c('socecon_lvl_cd','patient_distinct_socecon')], by = 'socecon_lvl_cd')

surgical_procedure_in_socecon_lvl <- surgical_procedure_in_socecon_lvl %>% mutate(
  perc_surgical_procedure_in_socecon_lvl = round((n_patient_surgical_socecon/patient_distinct_socecon)*100,2),
  united_surgical_procedure_in_socecon_lvl=paste0(n_patient_surgical_socecon, ' (',perc_surgical_procedure_in_socecon_lvl,'%)' )
)

## median time in socecon_lvl_cd 
mediantime_in_socecon_lvl <- dbGetQuery(conn=con,"
SELECT
	coalesce(socecon_lvl_cd,
	'Unknown')  as socecon_lvl_cd,
	median(surgical_procedure_dt -hospital_admission_dt) as median_time_adm_to_surgery,
	(QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt),
	0.75)-QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt),
	0.25)) as IQR_median_time_adm_to_surgery,
--round(mean(surgical_procedure_dt -hospital_admission_dt),2) as mean_time_adm_to_surgery
from
	main.patient_cohort
group by
	socecon_lvl_cd")

variables_in_hospital <- dbGetQuery(conn=con,"
SELECT
	coalesce(hospital_st,
	'Unknown')  as hospital_st,
count(distinct patient_id) as n_patient,
count(distinct hospitalization_episode_id) as n_hospitalization_episode_id,
count(distinct patient_id) filter (where surgical_procedure_bl = TRUE) as n_patient_with_surgery,
median(age_nm) as median_age, 
(QUANTILE_CONT(age_nm, 0.75)-QUANTILE_CONT(age_nm, 0.25)) as iqr_age,
count(distinct patient_id) filter (where sex_cd = '1') as men,
count(distinct patient_id) filter (where sex_cd = '2') as women,
median(surgical_procedure_dt -hospital_admission_dt) as median_time_adm_to_surgery,
(QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.75)-QUANTILE_CONT((surgical_procedure_dt -hospital_admission_dt), 0.25)) as IQR_median_time_adm_to_surgery,
count(distinct patient_id) filter (where exitus_bl = TRUE) as n_exitus,
from
	main.patient_cohort
group by
	hospital_st")

dbDisconnect(con, shutdown=TRUE)

```



## Descriptive


The following are shown 2 descriptive tables of the imported data:

- Table 1 shows the main characteristics of the patients registered in the database.

- Table 2 main characteristics of the patients and the medians from admission to surgery (in days) and their interquartile range grouped by hospital.



### Table 1: Descriptive inputs

```{r, echo=FALSE,warning=FALSE}
#| label: descriptive tables inputs1
descriptive_values_ <- melt(descriptive_values,id=NULL) 
anesthesia_type <- anesthesia_type %>% arrange(anesthesia_type_cd) %>%  dplyr::select(anesthesia_type_cd,united_anesthesia_type) %>% 
  mutate(variable = paste0('Anesthesia type: ',anesthesia_type_cd),
         value=united_anesthesia_type) %>% dplyr::select(variable,value)

comorbidities <- comorbidities %>% arrange(sum_comorbidities)  %>% dplyr::select(sum_comorbidities,united_comorbidities) %>% 
  mutate(variable = paste0('Nº comorbidities: ',sum_comorbidities),
         value=united_comorbidities) %>% dplyr::select(variable,value)

surgical_procedure <- surgical_procedure %>% arrange(surgical_procedure_cd) %>%  dplyr::select(surgical_procedure_cd,united_surgical_procedure) %>% 
  mutate(variable = paste0('Surgical procedure: ',surgical_procedure_cd),
         value=united_surgical_procedure) %>% dplyr::select(variable,value)

socecon_lvl <- socecon_lvl %>% arrange(socecon_lvl_cd) %>%  dplyr::select(socecon_lvl_cd,united_socecon) %>% 
  mutate(variable = paste0('Socioeconomic lvl: ',socecon_lvl_cd),
         value=united_socecon) %>% dplyr::select(variable,value)

descriptive_values_ <- rbind(descriptive_values_,anesthesia_type,comorbidities,surgical_procedure,socecon_lvl)
descriptive_values_$variable <- as.character(descriptive_values_$variable)


descriptive_values_$variable[descriptive_values_$variable %in% 'median_age'] <- 'Median of age'
descriptive_values_$variable[descriptive_values_$variable %in% 'iqr_age'] <- 'IQR of age'
descriptive_values_$variable[descriptive_values_$variable %in% 'n_patient'] <- 'Nº patients (unique)'
descriptive_values_$variable[descriptive_values_$variable %in% 'n_hospitalization_episode_id'] <- 'Nº hospitalization episode (unique)'
descriptive_values_$variable[descriptive_values_$variable %in% 'n_patient_with_surgery'] <- 'Nº patient with surgery (unique)'
descriptive_values_$variable[descriptive_values_$variable %in% 'median_time_adm_to_surgery'] <- 'Median time from admission to surgery (in days)'
descriptive_values_$variable[descriptive_values_$variable %in% 'IQR_median_time_adm_to_surgery'] <- 'IQR time from admission to surgery (in days)'
descriptive_values_$variable[descriptive_values_$variable %in% 'united_men'] <- 'Men'
descriptive_values_$variable[descriptive_values_$variable %in% 'united_women'] <- 'Women'


table <- descriptive_values_ %>% gt(rowname_col = "variable") %>% 
  tab_header(
    title = "Descriptive table input"
  ) %>% 
  tab_stubhead(label = "Variable") %>% 
     cols_label(
      variable = 'Variable',
      value = 'n (%)'
       ) %>% 
  tab_row_group(
    label = "Surgical procedure",
    rows = matches("surgical")) %>%
    tab_row_group(
    label = "Socioeconomic level",
    rows = matches("socioecon")) %>%
    tab_row_group(
    label = "Anesthesia",
    rows = matches("anesthesia")) %>%
    tab_row_group(
    label = "Comorbidities",
    rows = matches("comorbidities")) %>%
  row_group_order(groups = c(NA, "Socioeconomic level", "Surgical procedure", "Anesthesia","Comorbidities")) %>% 
    tab_style(
    style = cell_fill(color = "gray80"),
    locations = cells_row_groups(groups=everything())
  ) %>% 
  cols_align(
  align =  "center",
  columns = value)

div(style='height:600px; overflow-y: scroll', table)
rm(anesthesia_type,comorbidities,socecon_lvl,surgical_procedure,descriptive_values_,table)



```


### Table 2: Variables by hospital

```{r, echo=FALSE,warning=FALSE}
#| label: descriptive tables inputs2



table <- variables_in_hospital %>% gt() %>%
  tab_header(
    title = "Variables by hospital"
  ) %>%
     cols_label(
      hospital_st = 'Hospital',
      n_patient = 'Nº patients (unique)',
      n_hospitalization_episode_id = 'Nº hospitalization episode (unique)',
      n_patient_with_surgery = 'Nº patient with surgery (unique)',
      median_age = 'Median of age',
      iqr_age = 'IQR of age',
      men = 'Nº Men',
      women = 'Nº Women',
      median_time_adm_to_surgery = 'Median time from admission to surgery (in days)',
      IQR_median_time_adm_to_surgery = 'IQR time from admission to surgery (in days)',
      n_exitus = 'Nº exitus'
       ) %>%
        cols_align(
  align =  "center",
  columns = everything()) %>%
      tab_style(
    style = cell_text(align = "center"),
    locations = cells_row_groups())

div(style='height:600px; overflow-y: scroll', table)
  
rm(table)
```



## Analysis



```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: analysis

con = dbConnect(duckdb::duckdb(), dbdir=params$data_path, read_only=FALSE)

df_patient <- dbGetQuery(conn = con, "SELECT * FROM patient_cohort")
df_outpatient_prescription <- dbGetQuery(conn = con, "SELECT * FROM outpatient_prescription")
dbDisconnect(con, shutdown=TRUE)

df_patient <- df_patient %>% mutate_if(is.logical,as.numeric)


start_study <- as.Date('2017-01-01')
end_study <- as.Date('2023-12-31')

##### survival time: admission -> exitus in function to surgical procedure bl

df_patient <- df_patient %>% mutate(futime_adm_exi = case_when( 
                        !is.na(exitus_dt) ~ as.numeric(exitus_dt - hospital_admission_dt), 
                         is.na(exitus_dt) ~ as.numeric(end_study - hospital_admission_dt) ))
# 
# survobj <- Surv(time=df_patient$futime_adm_exi,
#                 event=df_patient$exitus_bl)
# 
# 
# 
# surv_fit <- survfit(survobj ~ surgical_procedure_bl, data=df_patient)



```

### Survival plot: Admission to exitus in fuction to have surgical procedure (Yes/No)

<!-- This analysis compares patients who have been treated with surgery versus those who have not received surgery during a follow-up time of 30 days from the date of admission to hospital.  -->

<!-- The Kaplan-Meier curve is shown below. -->

```{r, echo=FALSE,warning=FALSE}
#| label: survival plot

# ggsurvplot(surv_fit, data = df_patient, xlab="Days of follow-up",
#            ylab="Survival probability",
#            xlim=c(0,30),
#            ylim=c(0.6,1),
#            break.x.by=10,
#            conf.int = FALSE,
#            main="Product-Limit Survival Estimates", risk.table = TRUE)

```



```{r, echo=FALSE,warning=FALSE, output = FALSE}
#| label: analysis cont

df_patient_filter <- df_patient %>% filter(surgical_procedure_bl %in% 1)



df_patient_filter <- df_patient_filter %>% 
  mutate(
    futime_adm_surg = as.numeric(surgical_procedure_dt - hospital_admission_dt),
    futime_surg_end = case_when(
      !is.na(exitus_dt) ~ as.numeric(exitus_dt - surgical_procedure_dt),
      is.na(exitus_dt) ~ as.numeric(end_study - surgical_procedure_dt)),
    exitus_7days_bl = case_when(
      !is.na(exitus_dt) & futime_surg_end<= 7 ~ 1,
      TRUE ~ 0),
      exitus_30days_bl = case_when(
      !is.na(exitus_dt) & futime_surg_end<= 30 ~ 1,
      TRUE ~ 0),
    exitus_hosp_bl = case_when(
      hospital_discharge_dt==exitus_dt ~ 1,
      TRUE ~ 0)
  )

df_patient_filter <- df_patient_filter %>% mutate(
  period = (12*(year(surgical_procedure_dt)-year(start_study)) + month(surgical_procedure_dt)),
  weekday = weekdays(as.POSIXlt(surgical_procedure_dt),abbreviate = FALSE),
  day_month = day(surgical_procedure_dt)
)


df_model <- df_patient_filter %>% dplyr::select(exitus_hosp_bl,exitus_7days_bl,exitus_30days_bl,futime_adm_surg,age_nm,sex_cd,hospital_st,
                      socecon_lvl_cd,surgical_procedure_cd,anesthesia_type_cd,icu_admission_bl,anticoagulant_status_bl,antiagregant_status_bl,
                                              chronic_kidney_disease_bl,tobacco_copd_bl,
                                              depression_bl,serious_mental_illness_bl,alcohol_abuse_bl,             
                                              obesity_overweight_bl,osteoporosis_osteopenia_bl,dementia_bl,diabetes_bl,
                                              liver_disease_bl,pancreatic_disease_bl,inflamatory_bowel_disease_bl,rheumatology_disease_bl,
                                              spinal_cord_disease_injury_bl,serious_neurological_disease_bl,
                                              parkinson_huntintong_diseases_bl,seizure_disorder_convulsions_bl,congestive_heart_failure_bl,
                                              coronary_artery_disease_bl,cerebrovascular_disease_bl,peripheral_vascular_disease_bl,
                                              traumatic_brain_injury_bl,amputee_bl,period,weekday,day_month)

df_model$sex_cd <- as.character(df_model$sex_cd)
df_model$surgical_procedure_cd <- as.factor(df_model$surgical_procedure_cd)
df_model$hospital_st <- as.factor(df_model$hospital_st)
### select variables for model with 20% or less of missings

threshold = 0.2
df_model_aux <- df_model %>% select(!c(exitus_hosp_bl,exitus_7days_bl,exitus_30days_bl,period,day_month,futime_adm_surg))
a <- names(df_model_aux)[sapply(df_model_aux, function(x) (sum(is.na(x))/nrow(df_model_aux)) <= threshold)]
df_model_aux <- df_model_aux %>% select(a)
formula_model <- paste0(colnames(df_model_aux), collapse = ' + ')
rm(df_model_aux)

```


### Models

In order to study whether there is a relationship between exitus and the time elapsed between the date of admission and the date of surgery, three GAMM (Generalized Additive Mixed Models) models will be built following the same methodology but varying the dependent variable (y):

- Model 1: y = exitus in hospital (i.e. exitus date equals discharge from hospital date)

- Model 2: y = exitus in 7 days or less from the surgery date

- Model 3: y = exitus in 30 days or less from the surgery date


#### Simple models 

Two panels are shown for each of the three simple models according to the variable to be predicted (y):

- The first corresponds to the model plot with the y-axis being the model prediction with mean centred at zero and x-axis the main predictor variable, time from admission to surgery. 

<!-- - The second corresponds to the summary of the model. -->

- The second is an interactive summary table shows whether the probability of death is higher (red) or lower than average depending on the day after surgery. Thus, day 0-1 indicates that the patient is on the first day after surgery (first 24 hours), day 1-2 indicates that the patient is on the second day after surgery (between the first 24-48 hours), and so on. 

::: {.panel-tabset}

##### Plot model with variable to predict: exitus in hospital
```{r, echo=FALSE,warning=FALSE}
#| label: exitus in hospital simple plot

###### exitus in hospital #####


model_exitus_hosp_bl_simple <- mgcv::gam(exitus_hosp_bl ~ s(futime_adm_surg, bs = 'tp'), data = df_model, method = "REML")

# df_model[c('predict_model_exitus_hosp_bl_simple','se_predict_model_exitus_hosp_bl_simple')] <- model_exitus_hosp_bl_simple %>% predict(df_model, type = "response",se.fit=TRUE)
# 
# 
# 
# ggplot(data = df_model,  aes(x=futime_adm_surg, y=predict_model_exitus_hosp_bl_simple)) + 
#   geom_line() +
#   labs(title = '',
#            x = 'Time from admission to surgery',
#            y = 'Prediction to death in hospital') +
#   scale_x_continuous(breaks=seq(min(df_model$futime_adm_surg,na.rm=TRUE), max(df_model$futime_adm_surg,na.rm=TRUE), 2)) +
#   geom_hline(data=df_model, mapping=aes(yintercept=mean(predict_model_exitus_hosp_bl_simple,na.rm = TRUE)), color="red") +
#   theme(panel.background = element_blank(), axis.line = element_line(color='black')) 


plot(model_exitus_hosp_bl_simple, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)') 

p <- plot(model_exitus_hosp_bl_simple, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)') 
p1 <- data.frame(x=p[[1]][["x"]], y= p[[1]][["fit"]])
p1 <- p1 %>% 
  mutate(
    more_than_mean = ifelse(y>0,TRUE,FALSE),
    day = ceiling(x)
  )

p2_exitus_hosp_bl <- p1[!duplicated(p1$day),]

```


<!-- To annex -->

<!-- ##### Summary model with variable to predict: exitus in hospital -->

<!-- ```{r, echo=FALSE,warning=FALSE} -->
<!-- #| label: exitus in hospital simple summary -->

<!-- summary(model_exitus_hosp_bl_simple) -->


<!-- ``` -->


##### Summary table: More or less probability of exitus (depending on the day after surgery)

```{r, echo=FALSE,warning=FALSE}
#| label: exitus in hospital simple summary table
p2_exitus_hosp_bl$more_than_mean_descrp[p2_exitus_hosp_bl$more_than_mean == TRUE] <- 'More than mean'
p2_exitus_hosp_bl$more_than_mean_descrp[p2_exitus_hosp_bl$more_than_mean == FALSE] <- 'Less or equal than mean'


p2_exitus_hosp_bl$interval <- as.numeric(p2_exitus_hosp_bl$day) - 1
p2_exitus_hosp_bl <- p2_exitus_hosp_bl %>% filter(interval >= 0)
p2_exitus_hosp_bl$interval <- paste0(p2_exitus_hosp_bl$interval,' - ', p2_exitus_hosp_bl$day)

p2_exitus_hosp_bl %>% dplyr::select(interval,more_than_mean_descrp) %>% gt() %>%
  tab_header(
    title = ""
  ) %>%
     cols_label(
      interval = 'Day',
      more_than_mean_descrp = 'Difference with mean'
       ) %>%
  tab_style(
    style = cell_fill(color = "#F1553C"),
    locations = cells_body(
      columns = more_than_mean_descrp,
      rows = more_than_mean_descrp %in% 'More than mean'
    )) %>% 
        cols_align(
  align =  "center",
  columns = everything()) %>% 
    opt_interactive(use_filters = TRUE,
                  use_compact_mode = TRUE,
                  use_resizers = TRUE)

  

```


:::


::: {.panel-tabset}

##### Plot model with variable to predict: exitus in 7 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: exitus 7 days simple plot

###### exitus 7 days #####

model_exitus_7days_bl_simple <- mgcv::gam(exitus_7days_bl ~ s(futime_adm_surg, bs = 'tp'), data = df_model, method = "REML")

# df_model[c('predict_model_exitus_7days_bl_simple','se_predict_model_exitus_7days_bl_simple')] <- model_exitus_7days_bl_simple %>% predict(df_model, type = "response",se.fit=TRUE)
#
#
# ggplot(data = df_model,  aes(x=futime_adm_surg, y=predict_model_exitus_7days_bl_simple)) +
#   geom_line() +
#   labs(title = '',
#            x = 'Time from admission to surgery',
#            y = 'Prediction to death in 7 days after surgery') +
#     geom_hline(data=df_model, mapping=aes(yintercept=mean(predict_model_exitus_7days_bl_simple,na.rm = TRUE)), color="red") +
#     geom_vline(data=df_model, )
#   theme(panel.background = element_blank(), axis.line = element_rect(color='black'))

plot(model_exitus_7days_bl_simple, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')

p <- plot(model_exitus_7days_bl_simple, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p1 <- data.frame(x=p[[1]][["x"]], y= p[[1]][["fit"]])
p1 <- p1 %>%
  mutate(
    more_than_mean = ifelse(y>0,TRUE,FALSE),
    day = ceiling(x)
  )

p2_exitus_7days_bl <- p1[!duplicated(p1$day),]



```

<!-- To annex -->

<!-- ##### Summary model with variable to predict: exitus in hospital -->

<!-- ```{r, echo=FALSE,warning=FALSE} -->
<!-- #| label: exitus 7 days simple summary -->

<!-- summary(model_exitus_7days_bl_simple) -->


<!-- ``` -->


##### Summary table: More or less probability of exitus (depending on the day after surgery)

```{r, echo=FALSE,warning=FALSE}
#| label: exitus 7 days simple summary table
p2_exitus_7days_bl$more_than_mean_descrp[p2_exitus_7days_bl$more_than_mean == TRUE] <- 'More than mean'
p2_exitus_7days_bl$more_than_mean_descrp[p2_exitus_7days_bl$more_than_mean == FALSE] <- 'Less or equal than mean'


p2_exitus_7days_bl$interval <- as.numeric(p2_exitus_7days_bl$day) - 1
p2_exitus_7days_bl <- p2_exitus_7days_bl %>% filter(interval >= 0)
p2_exitus_7days_bl$interval <- paste0(p2_exitus_7days_bl$interval,' - ', p2_exitus_7days_bl$day)

p2_exitus_7days_bl %>% dplyr::select(interval,more_than_mean_descrp) %>% gt() %>%
  tab_header(
    title = ""
  ) %>%
     cols_label(
      interval = 'Day',
      more_than_mean_descrp = 'Difference with mean'
       ) %>%
  tab_style(
    style = cell_fill(color = "#F1553C"),
    locations = cells_body(
      columns = more_than_mean_descrp,
      rows = more_than_mean_descrp %in% 'More than mean'
    )) %>% 
        cols_align(
  align =  "center",
  columns = everything()) %>% 
    opt_interactive(use_filters = TRUE,
                  use_compact_mode = TRUE,
                  use_resizers = TRUE)

  

```


:::

::: {.panel-tabset}

##### Plot model with variable to predict: exitus in 30 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: exitus 30 days simple plot
###### exitus 30 days #####

#
# formula_model_exitus_30days_bl <- as.formula(paste0('exitus_30days_bl ~ s(futime_adm_surg) + s(futime_adm_surg, by = hospital_st) + s(period) + s(period, by = hospital_st) + s(day_month) + s(day_month, by = hospital_st) + ',formula_model))
#
#
# model_exitus_30days_bl <- mgcv::gam(formula_model_exitus_30days_bl, data = df_model, method = "REML")
# tab_model(model_exitus_30days_bl)

model_exitus_30days_bl_simple <- mgcv::gam(exitus_30days_bl ~ s(futime_adm_surg, bs = 'tp'), data = df_model, method = "REML")


# ggplot(data = df_model,  aes(x=futime_adm_surg, y=predict_model_exitus_30days_bl_simple)) +
#   geom_line() +
#   labs(title = '',
#            x = 'Time from admission to surgery',
#            y = 'Prediction to death in 30 days after surgery') +
#   theme(panel.background = element_blank(), axis.line = element_rect(color='black')) +
#   geom_hline(data=df_model, mapping=aes(yintercept=mean(predict_model_exitus_30days_bl_simple,na.rm = TRUE)), color="red")

plot(model_exitus_30days_bl_simple, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')

p <- plot(model_exitus_30days_bl_simple, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p1 <- data.frame(x=p[[1]][["x"]], y= p[[1]][["fit"]])
p1 <- p1 %>%
  mutate(
    more_than_mean = ifelse(y>0,TRUE,FALSE),
    day = ceiling(x)
  )

p2_exitus_30days_bl <- p1[!duplicated(p1$day),]


```




<!-- To annex -->

<!-- ##### Summary model with variable to predict: exitus in hospital -->

<!-- ```{r, echo=FALSE,warning=FALSE} -->
<!-- #| label: exitus 30 days simple summary -->

<!-- summary(model_exitus_30days_bl_simple) -->


<!-- ``` -->


##### Summary table: More or less probability of exitus (depending on the day after surgery)

```{r, echo=FALSE,warning=FALSE}
#| label: exitus 30 days simple summary table
p2_exitus_30days_bl$more_than_mean_descrp[p2_exitus_30days_bl$more_than_mean == TRUE] <- 'More than mean'
p2_exitus_30days_bl$more_than_mean_descrp[p2_exitus_30days_bl$more_than_mean == FALSE] <- 'Less or equal than mean'


p2_exitus_30days_bl$interval <- as.numeric(p2_exitus_30days_bl$day) - 1
p2_exitus_30days_bl <- p2_exitus_30days_bl %>% filter(interval >= 0)
p2_exitus_30days_bl$interval <- paste0(p2_exitus_30days_bl$interval,' - ', p2_exitus_30days_bl$day)

p2_exitus_30days_bl %>% dplyr::select(interval,more_than_mean_descrp) %>% gt() %>%
  tab_header(
    title = ""
  ) %>%
     cols_label(
      interval = 'Day',
      more_than_mean_descrp = 'Difference with mean'
       ) %>%
  tab_style(
    style = cell_fill(color = "#F1553C"),
    locations = cells_body(
      columns = more_than_mean_descrp,
      rows = more_than_mean_descrp %in% 'More than mean'
    )) %>% 
        cols_align(
  align =  "center",
  columns = everything()) %>% 
    opt_interactive(use_filters = TRUE,
                  use_compact_mode = TRUE,
                  use_resizers = TRUE)

  

```


:::


#### Model with all variables

::: {.panel-tabset}

##### Plot model with variable to predict: exitus in hospital


```{r, echo=FALSE,warning=FALSE}
#| label: exitus in hospital complex model plot

# sum_model_exitus_hosp_bl <- summary(model_exitus_hosp_bl)
# p_valor_nonspline <- as.data.frame(sum_model_exitus_hosp_bl$p.pv)
# p_valor_nonspline_signif <- p_valor_nonspline %>% filter(p_valor_nonspline$`sum_model_exitus_hosp_bl$p.pv` <= 0.05)
#
# p_valor_spline <- as.data.frame(sum_model_exitus_hosp_bl$s.table)
# p_valor_spline_signif <- p_valor_spline %>% filter(p_valor_spline$`p-value` <= 0.05)
#
# aux1 <- data.frame(variables_df = colnames(df_model))
# aux2 <- data.frame(variables_model = rownames(p_valor_nonspline_signif))
# aux3 <- data.frame(var=str_extract(aux2$variables_model,str_c(aux1$variables_df,collapse = '|')))
# aux3 <- aux3 %>% filter(!is.na(var))
# aux3 <- unique(aux3$var)
# formula_model <- paste0(aux3,collapse = '+')
# rm(aux1,aux2,aux3,p,p1)
#


### Filter by n_patient >=1200

variables_in_hospital_filter <- variables_in_hospital %>% filter(n_patient >= 1200)
df_model_filter <- df_model %>% filter(hospital_st %in% variables_in_hospital_filter$hospital_st)

formula_model_exitus_hosp_bl <- as.formula(paste0('exitus_hosp_bl ~ s(futime_adm_surg, bs = "tp") + s(futime_adm_surg, by = hospital_st, bs = "tp") + s(period, bs = "tp") + s(period, by = hospital_st, bs = "tp") + s(day_month, bs = "tp") + s(day_month, by = hospital_st, bs = "tp") + ',formula_model))

model_exitus_hosp_bl <- mgcv::gam(formula_model_exitus_hosp_bl, data = df_model_filter, method = "REML")


plot(model_exitus_hosp_bl, select = 1, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p <- plot(model_exitus_hosp_bl, select = 1, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p1 <- data.frame(x=p[[1]][["x"]], y= p[[1]][["fit"]])
p1 <- p1 %>%
  mutate(
    more_than_mean = ifelse(y>0,TRUE,FALSE),
    day = ceiling(x)
  )

p2 <- p1[!duplicated(p1$day),]

df_model_filter[c('predict_model_exitus_hospital_bl','se_predict_model_exitus_hospital_bl')] <- model_exitus_hosp_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)

output_model_exitus_hospital_bl <- df_model_filter %>% 
  group_by(hospital_st) %>% 
  summarise(mean_observed_model_exitus_hospital_bl =  mean(exitus_hosp_bl,na.rm=TRUE),
            sd_observed_model_exitus_hospital_bl = sd(exitus_hosp_bl,na.rm=TRUE),
            mean_predict_model_exitus_hospital_bl = mean(predict_model_exitus_hospital_bl,na.rm=TRUE),
            sd_predict_model_exitus_hospital_bl = sd(predict_model_exitus_hospital_bl,na.rm=TRUE),
            mean_residues_model_exitus_hospital_bl = mean(se_predict_model_exitus_hospital_bl,na.rm=TRUE),
            sd_residues_model_exitus_hospital_bl = sd(se_predict_model_exitus_hospital_bl,na.rm=TRUE),
            median_observed_model_exitus_hospital_bl =  median(exitus_hosp_bl,na.rm=TRUE),
            iqr_observed_model_exitus_hospital_bl = IQR(exitus_hosp_bl,na.rm=TRUE),
            median_predict_model_exitus_hospital_bl = median(predict_model_exitus_hospital_bl,na.rm=TRUE),
            iqr_predict_model_exitus_hospital_bl = IQR(predict_model_exitus_hospital_bl,na.rm=TRUE),
            median_residues_model_exitus_hospital_bl = median(se_predict_model_exitus_hospital_bl,na.rm=TRUE),
            iqr_residues_model_exitus_hospital_bl = IQR(se_predict_model_exitus_hospital_bl,na.rm=TRUE),
            )

```


<!-- ##### Summary model with variable to predict: exitus in hospital -->

<!-- ```{r, echo=FALSE,warning=FALSE} -->
<!-- #| label: exitus in hospital complex model summary -->

<!-- summary(model_exitus_hosp_bl) -->


<!-- ``` -->


##### Summary table: More or less probability of exitus (depending on the day after surgery)

```{r, echo=FALSE,warning=FALSE}
#| label: exitus in hospital complex summary table
p2_exitus_hosp_bl$more_than_mean_descrp[p2_exitus_hosp_bl$more_than_mean == TRUE] <- 'More than mean'
p2_exitus_hosp_bl$more_than_mean_descrp[p2_exitus_hosp_bl$more_than_mean == FALSE] <- 'Less or equal than mean'


p2_exitus_hosp_bl$interval <- as.numeric(p2_exitus_hosp_bl$day) - 1
p2_exitus_hosp_bl <- p2_exitus_hosp_bl %>% filter(interval >= 0)
p2_exitus_hosp_bl$interval <- paste0(p2_exitus_hosp_bl$interval,' - ', p2_exitus_hosp_bl$day)

p2_exitus_hosp_bl %>% dplyr::select(interval,more_than_mean_descrp) %>% gt() %>%
  tab_header(
    title = ""
  ) %>%
     cols_label(
      interval = 'Day',
      more_than_mean_descrp = 'Difference with mean'
       ) %>%
  tab_style(
    style = cell_fill(color = "#F1553C"),
    locations = cells_body(
      columns = more_than_mean_descrp,
      rows = more_than_mean_descrp %in% 'More than mean'
    )) %>% 
        cols_align(
  align =  "center",
  columns = everything()) %>% 
    opt_interactive(use_filters = TRUE,
                  use_compact_mode = TRUE,
                  use_resizers = TRUE)

  

```


:::


::: {.panel-tabset}

##### Plot model with variable to predict: exitus in 7 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: exitus 7 days complex model plot


formula_model_exitus_7days_bl <- as.formula(paste0('exitus_7days_bl ~ s(futime_adm_surg, bs = "tp") + s(futime_adm_surg, by = hospital_st, bs = "tp") + s(period, bs = "tp") + s(period, by = hospital_st, bs = "tp") + s(day_month, bs = "tp") + s(day_month, by = hospital_st, bs = "tp") + ',formula_model))

model_exitus_7days_bl <- mgcv::gam(formula_model_exitus_7days_bl, data = df_model_filter, method = "REML")

plot(model_exitus_7days_bl, select = 1, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p <- plot(model_exitus_7days_bl, select = 1, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p1 <- data.frame(x=p[[1]][["x"]], y= p[[1]][["fit"]])
p1 <- p1 %>%
  mutate(
    more_than_mean = ifelse(y>0,TRUE,FALSE),
    day = ceiling(x)
  )

p2_exitus_7days_bl <- p1[!duplicated(p1$day),]


df_model_filter[c('predict_model_exitus_7days_bl','se_predict_model_exitus_7days_bl')] <- model_exitus_7days_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)

output_model_exitus_7days_bl <- df_model_filter %>% 
  group_by(hospital_st) %>% 
  summarise(mean_observed_model_exitus_7days_bl =  mean(exitus_7days_bl,na.rm=TRUE),
            sd_observed_model_exitus_7days_bl = sd(exitus_7days_bl,na.rm=TRUE),
            mean_predict_model_exitus_7days_bl = mean(predict_model_exitus_7days_bl,na.rm=TRUE),
            sd_predict_model_exitus_7days_bl = sd(predict_model_exitus_7days_bl,na.rm=TRUE),
            mean_residues_model_exitus_7days_bl = mean(se_predict_model_exitus_7days_bl,na.rm=TRUE),
            sd_residues_model_exitus_7days_bl = sd(se_predict_model_exitus_7days_bl,na.rm=TRUE),
            median_observed_model_exitus_7days_bl =  median(exitus_7days_bl,na.rm=TRUE),
            iqr_observed_model_exitus_7days_bl = IQR(exitus_7days_bl,na.rm=TRUE),
            median_predict_model_exitus_7days_bl = median(predict_model_exitus_7days_bl,na.rm=TRUE),
            iqr_predict_model_exitus_7days_bl= IQR(predict_model_exitus_7days_bl,na.rm=TRUE),
            median_residues_model_exitus_7days_bl = median(se_predict_model_exitus_7days_bl,na.rm=TRUE),
            iqr_residues_model_exitus_7days_bl = IQR(se_predict_model_exitus_7days_bl,na.rm=TRUE),
            )

```

<!-- To annex -->

<!-- ##### Summary model with variable to predict: exitus in hospital -->

<!-- ```{r, echo=FALSE,warning=FALSE} -->
<!-- #| label: exitus 7 days simple summary -->

<!-- summary(model_exitus_7days_bl) -->


<!-- ``` -->


##### Summary table: More or less probability of exitus (depending on the day after surgery)

```{r, echo=FALSE,warning=FALSE}
#| label: exitus 7 days complex summary table
p2_exitus_7days_bl$more_than_mean_descrp[p2_exitus_7days_bl$more_than_mean == TRUE] <- 'More than mean'
p2_exitus_7days_bl$more_than_mean_descrp[p2_exitus_7days_bl$more_than_mean == FALSE] <- 'Less or equal than mean'


p2_exitus_7days_bl$interval <- as.numeric(p2_exitus_7days_bl$day) - 1
p2_exitus_7days_bl <- p2_exitus_7days_bl %>% filter(interval >= 0)
p2_exitus_7days_bl$interval <- paste0(p2_exitus_7days_bl$interval,' - ', p2_exitus_7days_bl$day)

p2_exitus_7days_bl %>% dplyr::select(interval,more_than_mean_descrp) %>% gt() %>%
  tab_header(
    title = ""
  ) %>%
     cols_label(
      interval = 'Day',
      more_than_mean_descrp = 'Difference with mean'
       ) %>%
  tab_style(
    style = cell_fill(color = "#F1553C"),
    locations = cells_body(
      columns = more_than_mean_descrp,
      rows = more_than_mean_descrp %in% 'More than mean'
    )) %>% 
        cols_align(
  align =  "center",
  columns = everything()) %>% 
    opt_interactive(use_filters = TRUE,
                  use_compact_mode = TRUE,
                  use_resizers = TRUE)

  

```


:::


::: {.panel-tabset}

##### Plot model with variable to predict: exitus in 30 days after surgery


```{r, echo=FALSE,warning=FALSE}
#| label: exitus 30 days complex plot
###### exitus 30 days #####

formula_model_exitus_30days_bl <- as.formula(paste0('exitus_30days_bl ~ s(futime_adm_surg, bs = "tp") + s(futime_adm_surg, by = hospital_st, bs = "tp") + s(period, bs = "tp") + s(period, by = hospital_st, bs = "tp") + s(day_month, bs = "tp") + s(day_month, by = hospital_st, bs = "tp") + ',formula_model))

model_exitus_30days_bl <- mgcv::gam(formula_model_exitus_30days_bl, data = df_model_filter, method = "REML")



plot(model_exitus_30days_bl, select = 1, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p <- plot(model_exitus_30days_bl, select = 1, xlab = 'Time from admission to surgery (days)', ylab = 'prediction exitus in hospital - mean(prediction exitus in hospital)')
p1 <- data.frame(x=p[[1]][["x"]], y= p[[1]][["fit"]])
p1 <- p1 %>%
  mutate(
    more_than_mean = ifelse(y>0,TRUE,FALSE),
    day = ceiling(x)
  )

p2_exitus_30days_bl <- p1[!duplicated(p1$day),]

df_model_filter[c('predict_model_exitus_30days_bl','se_predict_model_exitus_30days_bl')] <- model_exitus_30days_bl %>% predict(df_model_filter, type = "response",se.fit=TRUE)

output_model_exitus_30days_bl <- df_model_filter %>% 
  group_by(hospital_st) %>% 
  summarise(mean_observed_model_exitus_30days_bl =  mean(exitus_30days_bl,na.rm=TRUE),
            sd_observed_model_exitus_30days_bl = sd(exitus_30days_bl,na.rm=TRUE),
            mean_predict_model_exitus_30days_bl = mean(predict_model_exitus_30days_bl,na.rm=TRUE),
            sd_predict_model_exitus_30days_bl = sd(predict_model_exitus_30days_bl,na.rm=TRUE),
            mean_residues_model_exitus_30days_bl = mean(se_predict_model_exitus_30days_bl,na.rm=TRUE),
            sd_residues_model_exitus_30days_bl = sd(se_predict_model_exitus_30days_bl,na.rm=TRUE),
            median_observed_model_exitus_30days_bl =  median(exitus_30days_bl,na.rm=TRUE),
            iqr_observed_model_exitus_30days_bl = IQR(exitus_30days_bl,na.rm=TRUE),
            median_predict_model_exitus_30days_bl = median(predict_model_exitus_30days_bl,na.rm=TRUE),
            iqr_predict_model_exitus_30days_bl= IQR(predict_model_exitus_30days_bl,na.rm=TRUE),
            median_residues_model_exitus_30days_bl = median(se_predict_model_exitus_30days_bl,na.rm=TRUE),
            iqr_residues_model_exitus_30days_bl = IQR(se_predict_model_exitus_30days_bl,na.rm=TRUE),
            )


```




<!-- To annex -->

<!-- ##### Summary model with variable to predict: exitus in hospital -->

<!-- ```{r, echo=FALSE,warning=FALSE} -->
<!-- #| label: exitus 30 days simple summary -->

<!-- summary(model_exitus_30days_bl) -->


<!-- ``` -->


##### Summary table: More or less probability of exitus (depending on the day after surgery)

```{r, echo=FALSE,warning=FALSE}
#| label: exitus 30 days complex summary table
p2_exitus_30days_bl$more_than_mean_descrp[p2_exitus_30days_bl$more_than_mean == TRUE] <- 'More than mean'
p2_exitus_30days_bl$more_than_mean_descrp[p2_exitus_30days_bl$more_than_mean == FALSE] <- 'Less or equal than mean'


p2_exitus_30days_bl$interval <- as.numeric(p2_exitus_30days_bl$day) - 1
p2_exitus_30days_bl <- p2_exitus_30days_bl %>% filter(interval >= 0)
p2_exitus_30days_bl$interval <- paste0(p2_exitus_30days_bl$interval,' - ', p2_exitus_30days_bl$day)

p2_exitus_30days_bl %>% dplyr::select(interval,more_than_mean_descrp) %>% gt() %>%
  tab_header(
    title = ""
  ) %>%
     cols_label(
      interval = 'Day',
      more_than_mean_descrp = 'Difference with mean'
       ) %>%
  tab_style(
    style = cell_fill(color = "#F1553C"),
    locations = cells_body(
      columns = more_than_mean_descrp,
      rows = more_than_mean_descrp %in% 'More than mean'
    )) %>% 
        cols_align(
  align =  "center",
  columns = everything()) %>% 
    opt_interactive(use_filters = TRUE,
                  use_compact_mode = TRUE,
                  use_resizers = TRUE)

  

```


:::


